### 2018-07-09 src/EC50aggregation.Rmd
publishing orphan commit

---
title: "EC50 calculation + calculation and upload"
author: "Andreas Scharmüller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
editor_options:
  chunk_output_type: console
---

The following script collects different ecotoxicological tests and extracts EC50 values for multiple organisms for different pesticide compounds. The compounds found in the __psm-data.frame__ are compiled in the EC50sensitivity_data.Rmd script and loaded here in a first step. As a first source for ecotoxicological test data the EPA ECOTOX data base is queried (data-epa). The data itself is stored in a PostgreSQL data base built accoring to [this guide](http://edild.github.io/localecotox/). Secondly the bfg_monitoring data-base is queried specifically for values of __Daphnia magna__ tests. The vast amount of test therein was collected by Eduard Szöcs. In a third step data from the [PPDB](https://sitem.herts.ac.uk/aeru/ppdb/en/search.htm) is queried. This is not in accordance to their data policy, but it works. After quering the different data bases, habitat information on family level which was compiled by Verena, Gesa and me is added to the test data set (epa1). The acquire test data is then aggregated (min, percentile, mean) for single substances (aggregate-ec50-all) over all organisms and over freshwater organisms only (aggregate-ec50-freshwater). In a last step the aggregated test data is uploaded to a PostgreSQL data base.

# Setup

```{r setup}
#rm(list = ls())
# switches
online = FALSE
local = FALSE # have you built the EPA ECOTOX DB locally?
plots = FALSE

# packages
require(data.table)
require(RPostgreSQL)
require(taxize)
require(webchem)
require(ggplot2)
require(ggrepel)
require(vegan)

# variables
cachedir = '/home/andreas/Documents/Projects/species-sensitivity/cache'
fundir = '/home/andreas/Documents/Projects/species-sensitivity/functions'
plotdir = '/home/andreas/Documents/Projects/species-sensitivity/plots'
srcdir = '/home/andreas/Documents/Projects/species-sensitivity/src'

# source
source('/home/andreas/Documents/UBA/Project/R/src/DB_access.R')
source(file.path(fundir, 'ppdb-0-0-5.R'))
source(file.path(fundir, 'ppdb_own_functions.R'))
source(file.path(fundir, 'casconv.R'))
source(file.path(fundir, 'functions.R')) # contains gm_mean(), ksource()
source(file.path(fundir, 'agg_group.R'))
source(file.path(fundir, 'plot_outlier.R'))
```

# Load substances
```{r subst1}
## load data
psm = readRDS(file.path(cachedir, 'psm1.rds')) # created in EC50sensitivity_data.Rmd
psm = psm[ !is.na(casnr) ] # only metabolite and sumed pesticides
psm = psm[ !subst_name %in% c('Desethylsimazin') ] # casnr: 1007289; duplicated with Desethylsimazin

## for Nadin and verena
# psm = fread('/home/andreas/Documents/data/romania/det_compounds_with_CAS.csv')
# psm[ , casnr := gsub('-', '', cas) ]
# psm[ , `:=`
#      (variable_id = NA,
#       subst_name = NA,
#       psm_type = NA)]

## CHECKs
cas_check = 
  psm[ is.na(casnr) | casnr == '' | 
       is.na(cas) | cas == '' ]

if (nrow(cas_check) != 0) {
  warning(nrow(cas_check), 'missing CAS or CASNR entries.')
}

rm(cas_check)
```

# PubChem
```{r cas-resolver}
## pubchem
cas = psm$cas

if (online) {

  cid = get_cid(cas)

  pc_l = list()
  for (i in 1:length(cid)) {
    cas1 = names(cid[i])
    cid1 = cid[[i]]
    message('Querying: CAS:', cas1, '; CID:', cid1)

    pc1 = pc_prop(cid1)

    pc_l[[i]] = pc1
    names(pc_l)[i] = cas1
  }

  saveRDS(pc_l, file.path(cachedir, 'pc_l.rds'))
  saveRDS(cid, file.path(cachedir, 'cid.rds'))

} else {
  pc_l = readRDS(file.path(cachedir, 'pc_l.rds'))
  cir = readRDS(file.path(cachedir, 'cid.rds'))
}

# convert NAs from negative results to data.tables to be able to bind 'em
pc_l[ is.na(pc_l) ] = lapply(pc_l[ is.na(pc_l) ], data.table)

pubchem = rbindlist(pc_l, fill = TRUE, idcol = 'cas')

# cleaning
rm(cir)

# how to handle duplicated CAS
# test = pubchem[ , .(N = .N,
#              InChI = paste0(InChIKey),
#              IUPACName), by = cas][ N > 1]
```

As some CAS are duplicated the Pubchem data base is queried to get InChikeys and IUPAC names to be able to identify unique substances.
```{r subst2}
psm[pubchem, on = 'cas', iupac_name := i.IUPACName ]
# replacing missing subst_name with IUPAC names
psm[ , subst_name := ifelse(is.na(subst_name), iupac_name, subst_name)]
saveRDS(psm, file.path(cachedir, 'psm2.rds'))
```

# Ecotox test data

## Source EPA ECOTOX data
```{r SOURCE-epa}
ksource(file.path(srcdir, 'source_epa.Rmd'))
```

## Source bfg_monitoring data
Not needed anymore.

## Source PPDB data
```{r SOURCE-ppdb}
ksource(file.path(srcdir, 'source_ppdb.Rmd'))
# data cleaning
ppdb_dt[ latin_BIname %in% c('A4', 'A5', 'Cyanobacteria', 'Cyprinidae', 'Green algae', 'P4'),
         latin_BIname := 'Unknown species' ]
```

## Source Malaj data
```{r SOURCE-mlaj}
ksource(file.path(srcdir, 'source_malaj.Rmd'))
```

## Bind test reults
```{r tests-bind}
# bind the results from different sources
tests = rbindlist(list(epa2, ppdb_dt, malaj_m), fill = TRUE) # dm, 
```

## Errata
```{r tests-errata}
tests[ latin_BIname == 'Notropis ludibundis', `:=`
       (latin_BIname = 'Notropis ludibundus',
        latin_short = 'N. ludibundus')]

## remove hybrid X marking
tests[ , latin_BIname := gsub('\\sX\\s', ' ', latin_BIname) ]
## remove sp.
tests[ , latin_BIname := gsub('\\ssp\\.', '', latin_BIname) ]
## change species names
tests[ latin_BIname == 'Paratelphusa masoniana', latin_BIname := 'Parathelphusa masoniana' ]
tests[ latin_BIname == 'Tresus nuttalli', latin_BIname := 'Tresus nuttallii' ]
tests[ latin_BIname == 'Mysidopsis bahia', latin_BIname := 'Americamysis bahia' ]
tests[ latin_BIname == 'Planaria gonocephala', latin_BIname := 'Dugesia gonocephala' ]

## change family names
tests[ latin_BIname == 'Tubifex', family_epa := 'Naididae' ]
## change species and family names
tests[ latin_BIname == 'Chlorella fusca', `:=`
       (latin_BIname = 'Scenedesmus fuscus',
         family_epa = 'Scenedesmaceae ') ] # according to worms

## save the results locally
saveRDS(tests, file.path(cachedir, 'tests1.rds'))
```

# Taxonomic data

```{r MERGE-taxonomic-data}
#### Taxonomic data ----
ksource(file.path(srcdir, 'speciesinfo.Rmd'))

# Merge
tests = merge(tests, tx_dt, on = 'latin_BIname', all = TRUE)

#### Check taxize family classification ----
# tests[ source == 'epa_ecotox' & family != family_epa,
#        c('latin_BIname', 'family', 'family_epa', 'source_tax') ][ , .N, .(latin_BIname, family, family_epa, source_tax)][order(-N)]
use_epa_fam = c('Lemna minor', 'Lemna gibba', 'Lemna aequinoctialis', 'Viviparus bengalensis', 'Rattus rattus', 'Phytophthora capsici', 'Ganaspidium utilis', 'Indonaia caerulea', 'Phytophthora citricola', 'Rutilus rutilus', 'Sigara substriata', 'Stenonema candidum', 'Eriophyes datura', 'Hyalella curvispina', 'Neopanope texana', 'Paecilomyces farinosus', 'Melanopsis dufouri', 'Thiara lineata', 'Phytophthora cactorum', 'Waitea circinata', 'Culex melanurus', 'Lithodes antarcticus', 'Meloidogyne incognita', 'Paratelphusa masoniana', 'Physalaemus biligonigerus', 'Stethorus japonicus', 'Aphanizomenon flos-aquae', 'Caridina rajadhari', 'Palaemonetes africanus', 'Plea atomaria', 'Anacystis alpicola', 'Anisops bouvieri', 'Clibanarius africanus', 'Katelysia opima', 'Palustra laboulbeni', 'Physalaemus centralis', 'Tresus nuttalli', 'Anticarsia gemmatalis', 'Elasmopus bampo', 'Physalaemus santafecinus', 'Planaria gonocephala', 'Podisus nigrispinus', 'Ranatra elongata', 'Zonotrichia albicollis', 'Oziotelphusa senex',
'Perca', 'Nitzschia', 'Lemna', 'Libellula', 'Opius', 'Hydra', 'Asellus', 'Cyclops', 'Cypria', 'Sphaerium', 'Crangon', 'Danio', 'Lucifer', 'Micronecta', 'Sagitta', 'Alona', 'Ameletus', 'Amphora', 'Colpoda', 'Lestes', 'Micropterus', 'Physella', 'Ulnaria', 'Choristoneura rosaceana', 'Cydia pomonella', 'Epiphyas postvittana', 'Helicoverpa armigera', 'Pholetesor ornigis', 'Phthorimaea operculella', 'Plodia interpunctella', 'Plutella xylostella', 'Sclerotinia sclerotiorum', 'Spodoptera littoralis')

#### final family classification ----
## by taxize wrongly classified columns
tests[ , family := ifelse(latin_BIname %in% use_epa_fam, family_epa, family_tax) ]

## family errata
tests[ latin_BIname == 'Pseudokirchneriella subcapitata',
       family := 'Selenastraceae']

#! INTERMEDIATE: They are not yet classified in speciesinfo.org
#! were added when time span was expanded to 360h
tests[ source == 'epa_ecotox' & is.na(family),
       family := 'family_epa' ]

# save the results locally
saveRDS(tests, file.path(cachedir, 'tests2.rds'))
```

# Habitat data
```{r MERGE-habitat-data}
#### Habitat data ----
# self compiled list
ksource(file.path(srcdir, 'familylookup.Rmd'))
# worms
ksource(file.path(srcdir, 'habitat_classification.Rmd'))
# freshwaterecology.info
#! todo
# algaebase
#! scraping is against their terms of use

## Merge
tests = merge(tests, lookup, by = 'family', all.x = TRUE)
tests = merge(tests, lookup_worms_sp, by = 'latin_BIname', all.x = TRUE)

#### Create isFre final column ----
# Freshwater
tests[ , 
       isFre_sum := rowSums(sapply(.SD, as.numeric), na.rm = TRUE),
       .SDcols = c('isFre', 'isFre_w') ]
tests[ , isFre_fin := ifelse(isFre_sum >= 1, '1', '0') ][ , isFre_sum := NULL ]
# Marine
tests[ , 
       isMar_sum := rowSums(sapply(.SD, as.numeric), na.rm = TRUE),
       .SDcols = c('isMar', 'isMar_w') ]
tests[ , isMar_fin := ifelse(isMar_sum >= 1, '1', '0') ][ , isMar_sum := NULL ]
# Terrestrial
tests[ , 
       isTer_sum := rowSums(sapply(.SD, as.numeric), na.rm = TRUE),
       .SDcols = c('isTer', 'isTer_w') ]
tests[ , isTer_fin := ifelse(isTer_sum >= 1, '1', '0') ][ , isTer_sum := NULL ]

#### Classify Invertebrates ----
#invertebrates_phylum = c("Annelida", "Bryozoa", "Chaetognatha", "Ciliophora", "Cnidaria", "Gastrotricha", "Echinodermata", "Mollusca", "Nematoda", "Nemertea", "Platyhelminthes", "Porifera", "Rotifera")
inv_makro_phylum = c('Annelida', 'Echinodermata', 'Mollusca', 'Nemertea', 'Platyhelminthes', 'Porifera')

inv_mikro_phylum = c('Bryozoa', 'Chaetognatha', 'Ciliophora', 'Cnidaria', 'Gastrotricha', 'Nematoda', 'Rotifera')

#invertebrates_subphylum = c('Crustacea') # phylum: Arthropoda, Chordata
inv_makro_subphylum = c('Crustacea')
inv_makro_class = c('Arachnida', 'Diplopoda', 'Entognatha', 'Insecta') # phylum: Arthropoda

invertebrates_makro = c(inv_makro_phylum, inv_makro_subphylum, inv_makro_class)
invertebrates_mikro = inv_mikro_phylum

tests[ , supgroup2 := ifelse(supgroup %in% invertebrates_makro, 'Makro_Inv',
                      ifelse(supgroup %in% invertebrates_mikro, 'Mikro_Inv', supgroup)) ]

#### Classify trophic level
autotrophs = c('Plants', 'Algae', 'Bryophyta')
tests[ , trophic_lvl := ifelse(supgroup2 %in% autotrophs,
                               'autotrophic', 'heterotrophic') ]
# cleaning
rm(list = grep('inv_', ls(), value = TRUE))
```

# Checks and save

Check whether al entries have CAS, CASNR and substance and species names.
```{r checks}
subst_check = 
  tests[ is.na(casnr) | casnr == '' |
         is.na(cas) | cas == '' |
         is.na(subst_name) | subst_name == '' ]
tax_check = 
  tests[ is.na(latin_BIname) | latin_BIname == '' |
         is.na(group) | group == '' |
         is.na(supgroup) | supgroup == '' |
         is.na(supgroup2) | supgroup == '' ]
value_check = 
  tests[ is.na(value) | value == '' ]
fresh_check =
  tests[ is.na(isFre_fin) | isFre_fin == '' ]
marin_check = 
  tests[ is.na(isMar_fin) | isMar_fin == '' ]
terre_check = 
  tests[ is.na(isTer_fin) | isTer_fin == '' ]

if (nrow(subst_check) != 0) {
  warning(nrow(subst_check), ' missing CAS, CASNR or substance names.')
}
if (nrow(tax_check) != 0)  {
  warning(nrow(tax_check), ' missing latin_BIname, group, supgroup or supgroup2 entry.')
}
if (nrow(value_check) != 0) {
  warning(nrow(value_check), ' missing value entries.')
}
if (nrow(fresh_check) != 0) {
  warning(nrow(value_check), ' missing freshwater classifications.')
}
if (nrow(marin_check) != 0) {
  warning(nrow(marin_check), ' missing marine classifications.')
}
if (nrow(terre_check) != 0) {
  warning(nrow(terre_check), ' missing terrestrial classifications.')
}

# cleaning
rm(list = grep('check', ls(), value = TRUE))

# write to cache
saveRDS(tests, file.path(cachedir, 'tests3.rds'))
```

# Overview table

```{r overview-table}
taxa_fresh = tests[ isFre_fin == '1' , .N, .(supgroup2, supgroup, group) ][order(supgroup2, supgroup, group)]
taxa_marin = tests[ isMar_fin == '1' , .N, .(supgroup2, supgroup, group) ][order(supgroup2, supgroup, group)]
taxa_terre = tests[ isTer_fin == '1' , .N, .(supgroup2, supgroup, group) ][order(supgroup2, supgroup, group)]

taxa_fresh
taxa_marin
taxa_terre

# cleaning
rm(list = grep('taxa', ls(), value = TRUE))
```

# EC50 aggregation

```{r read-tests-data}
tests3 = readRDS(file.path(cachedir, 'tests3.rds'))
tests3[ str_count(tests3$latin_BIname, '\\S+') < 2, 
        latin_BIname := paste0(latin_BIname, ' sp.') ]
```

## Malaj aggregate
```{r AGGREGATE-malaj}
# aggregating by casnr
malaj_agg =
  tests3[ source == 'malaj',
          j = .(mj50f_al4896_vl = min(.SD[latin_BIname == 'Pseudokirchneriella subcapitata']$value),
                mj50f_da48_vl = min(.SD[latin_BIname == 'Daphnia magna']$value),
                mj50f_fi96_vl = min(.SD[latin_BIname == 'Pimephales promelas']$value)),
          by = .(casnr) ]

# Inf to NA
for (i in names(malaj_agg)) {
  malaj_agg[get(i) == Inf, (i) := NA]
}

# add info column
malaj_agg[ , mj50f_al4896_tax := ifelse(!is.na(mj50f_al4896_vl), 'Pseudokirchneriella subcapitata', NA) ]
malaj_agg[ , mj50f_da48_tax := ifelse(!is.na(mj50f_da48_vl), 'Daphnia magna', NA) ]
malaj_agg[ , mj50f_fi96_tax := ifelse(!is.na(mj50f_fi96_vl), 'Pimephales promelas', NA) ]

# add test count column (careful, some are estimated!)
malaj_agg[ , mj50f_al4896_n := ifelse(!is.na(mj50f_al4896_vl), 1, NA) ]
malaj_agg[ , mj50f_da48_n := ifelse(!is.na(mj50f_da48_vl), 1, NA) ]
malaj_agg[ , mj50f_fi96_n := ifelse(!is.na(mj50f_fi96_vl), 1, NA) ]

# colorder
setcolorder(malaj_agg, c('casnr',
                         'mj50f_al4896_vl', 'mj50f_al4896_tax', 'mj50f_al4896_n',
                         'mj50f_da48_vl', 'mj50f_da48_tax', 'mj50f_da48_n',
                         'mj50f_fi96_vl', 'mj50f_fi96_tax', 'mj50f_fi96_n'))
```

## PPDB aggregate
```{r AGGREGATE-ppdb}
ppdb_agg = 
  tests3[ source == 'ppdb',
          j = .(# algae
                pp50f_al72_vl = min(.SD[ppdb_group == 'algae']$value),
                pp50f_al72_tax = .SD[ppdb_group == 'algae' & which.min(value)]$latin_BIname,
                # aquatic plants
                pp50f_pl168_vl = min(.SD[ppdb_group == 'aquatic plants']$value),
                pp50f_pl168_tax = .SD[ppdb_group == 'aquatic plants' & which.min(value)]$latin_BIname,
                # Daphnia
                pp50f_da48_vl = min(.SD[latin_BIname %like% '(?i)daphnia']$value),
                pp50f_da48_tax = .SD[latin_BIname %like% '(?i)daphnia' & which.min(value)]$latin_BIname,
                # crustaceans
                pp50f_cr96_vl = min(.SD[ppdb_group == 'aquatic crustaceans']$value),
                pp50f_cr96_tax = .SD[ppdb_group == 'aquatic crustaceans' & which.min(value)]$latin_BIname,
                # sediment dwelling organisms (mostly Chironomidae)
                pp50f_ch96_vl = min(.SD[ppdb_group == 'sediment dwelling organisms']$value),
                pp50f_ch96_tax = .SD[ppdb_group == 'sediment dwelling organisms' & which.min(value)]$latin_BIname,
                # fish
                pp50f_fi96_vl = min(.SD[ppdb_group == 'fish']$value),
                pp50f_fi96_tax = .SD[ppdb_group == 'fish' & which.min(value)]$latin_BIname),
          by = .(casnr) ]

# cleaning
for (i in names(ppdb_agg)) {
  ppdb_agg[get(i) == Inf, (i) := NA]
}

# add count column
ppdb_agg[ , pp50f_al72_n := ifelse(!is.na(pp50f_al72_vl), 1, NA) ]
ppdb_agg[ , pp50f_pl168_n := ifelse(!is.na(pp50f_pl168_vl), 1, NA) ]
ppdb_agg[ , pp50f_da48_n := ifelse(!is.na(pp50f_da48_vl), 1, NA) ]
ppdb_agg[ , pp50f_cr96_n := ifelse(!is.na(pp50f_cr96_vl), 1, NA) ]
ppdb_agg[ , pp50f_ch96_n := ifelse(!is.na(pp50f_ch96_vl), 1, NA) ]
ppdb_agg[ , pp50f_fi96_n := ifelse(!is.na(pp50f_fi96_vl), 1, NA) ]

#### Combine Invertebrates ----
ppdb_agg_m = 
  melt(ppdb_agg, id.vars = 'casnr',
  measure.vars = list(c('pp50f_da48_vl', 'pp50f_cr96_vl', 'pp50f_ch96_vl'),
                      c('pp50f_da48_tax', 'pp50f_cr96_tax', 'pp50f_ch96_tax'),
                      c('pp50f_da48_n', 'pp50f_cr96_n', 'pp50f_ch96_n')))
ppdb_agg_m_agg =
  ppdb_agg_m[ !is.na(value1) & !is.infinite(value1),
              .(pp50f_iv4896_min = min(value1),
                pp50f_iv4896_tax = .SD[which.min(value1)]$value2,
                pp50f_iv4896_info = paste0(.SD[order(value1)]$value2, ' (',
                                           .SD[order(value1)]$value1, 'ug/L)',
                                           collapse = ' - '),
                pp50f_iv4896_n = sum(.SD[!is.na(value1)]$value3)),
              casnr ]

ppdb_agg = ppdb_agg[ppdb_agg_m_agg, on = 'casnr']
# removing Inf and (µg/L)
for (i in names(ppdb_agg)) {
  ppdb_agg[get(i) == ' (ug/L)', (i) := NA]
}

# col order
names(ppdb_agg)
setcolorder(ppdb_agg, c('casnr',
                        'pp50f_al72_vl', 'pp50f_al72_tax', 'pp50f_al72_n',
                        'pp50f_pl168_vl', 'pp50f_pl168_tax', 'pp50f_pl168_n',
                        'pp50f_da48_vl', 'pp50f_da48_tax', 'pp50f_da48_n',
                        'pp50f_cr96_vl', 'pp50f_cr96_tax', 'pp50f_cr96_n',
                        'pp50f_ch96_vl', 'pp50f_ch96_tax', 'pp50f_ch96_n',
                        'pp50f_fi96_vl', 'pp50f_fi96_tax', 'pp50f_fi96_n',
                        'pp50f_iv4896_min', 'pp50f_iv4896_tax', 'pp50f_iv4896_info', 'pp50f_iv4896_n'))
```

### Malaj-PPDB check
```{r malaj-ppdb-check}
# check
malj_check = length(which(duplicated(malaj_agg$casnr)))
ppdb_check = length(which(duplicated(ppdb_agg$casnr)))

if (malj_check != 0) {
  stop(malj_check, ' duplicated CAS entries.')
}
if (ppdb_check != 0) {
  stop(ppdb_check, ' duplicated CAS entries.')
}

# cleaning
rm(malj_check, ppdb_check)
```

## EPA aggregate
```{r AGGREGATE-epa1, eval=FALSE}
# ksource(file.path(srcdir, 'EC50aggregation_epa1.Rmd'))
```

## EPA aggregate 2
```{r AGGREGATE-epa2}
ksource(file.path(srcdir, 'EC50aggregation_epa2.Rmd'))
```

# Merge
```{r MERGE-ec50}
source_l = list(malaj_agg, ppdb_agg, ep50f_agg)
ec50 = Reduce(function(...) merge(..., on = 'casnr', all = TRUE), source_l)
ec50[tests3, on = 'casnr', subst_name := i.subst_name]
setcolorder(ec50, c(1, ncol(ec50), 3:ncol(ec50)-1))
```

## Check
```{r ec50-check}
na_check = ec50[ is.na(casnr) | is.na(subst_name) ]
if (nrow(na_check) != 0) {
  warning(nrow(na_check), ' missing CAS, CASNR or substance names.')
}
rm(na_check)
```

# Write to database
```{r disclaimer}
su = length(unique(tests3$casnr)) # substances
sp = length(as.character(unique(tests3$latin_BIname))) # species
fa = length(unique(tests3$family)) # families
ha = nrow(lookup)
```

```{r WRITE-postgres, eval=FALSE}
drv = dbDriver("PostgreSQL")
con = dbConnect(drv, user = DBuser, dbname = DBname_1, host = DBhost, port = DBport, password = DBpassword)

schema = 'phch'
table = 'phch_ec50'

## upload
dbSendQuery(con, paste0("DROP TABLE IF EXISTS  ", schema, ".", table, ";"))
dbWriteTable(con, ec50,
             name = c(schema, table), row.names = FALSE)

## User rights
dbSendQuery(con, paste0("GRANT ALL ON TABLE ", schema, ".", table, " TO scharmueller;
                  GRANT ALL ON TABLE ", schema, ".", table, " TO bfg;
                  GRANT SELECT ON TABLE ", schema, ".", table, " TO bfg_read;"))

## Primary Key
# dbSendQuery(con, paste0("ALTER TABLE ", schema, ".", table, " ADD PRIMARY KEY (variable_id);"))

## COMMENT ON
dbSendQuery(con, paste0("COMMENT ON TABLE  ", schema, ".", table, "  IS '
pesticide ec50 table fo specific organism groups per CASNR:

#### Malaj et al. (2014) data ----
(1) mj50_<ps|dm|pp> = EC50 values from Malaj for dm-Daphnia magna, pp-Pimephales promelas, ps-Pseudokirchneriella subcapitata (Already aggregated).
(2) mj50_<dm_type|pp_type|ps_type> = value type used by Malaj et al.:
(B-Baseline toxicity, E-Empirical, P1-4-Predictet(3,4-trustworthy;1,2-not))

#### PPDB data ----
(1) pp50f_<al|pl|da|cr|ch|fi>
EC50 vlaues from PPDB for iv-Invertebrates, al-Algae and fi-Fish (the minimum per CASNR and tax. group was taken).
(2) pp50f_<al|pl|da|cr|ch|fi>_info
string containing all EC50 values per CASNR and tax. group.

#### EPA data ----
(1) n<f|m|t|n>_<al|pl|iv|da|cr|ch|in|fi|am><48|72|96|168>
number of tests for specific substance and organism group (here: algae).
(2) ep50<f|m|t|n>_<al|pl|iv|da|cr|ch|in|fi|am><48|72|96|168>_<min|gm|mn>
e.g:
ep50_al72_min = ec50_algae72h_minimum_value
ep50_da48_gm = ec50_Daphniidae48h_geometric mean (with deleted ouliers)
ep50f_al4896_mn = ec50freshwater_algae48h-96h_mean
! For those w/o time specification tests with a duration from 24 to 120h (for plants: 360h) were used.
(3) ep50<f|m|t|n>_<al|pl|iv|da|cr|ch|in|fi|am><48|72|96|168>_info
info column which taxa wered used (orderd by value), the EPA ecotox ref. number and the value (separated by comma)
e.g Daphnia magna-(984)-50ugL,Daphnia magna-(984)-50ugL
(4) ep50<f|m|t|n>_<al|pl|iv|da|cr|ch|in|fi|am><48|72|96|168>_out
colum which denotes test results that were excluded based on outlier detection

Aggregated EPA taxa:
", freshw_grp[ ,1], "
", freshw_grp[ ,2], "
", freshw_grp[ ,3], "
", freshw_grp[ ,4], "

---
Abbreviations:
value types:
vl = value (there was only one value, hence no aggregation)
min = minimum
gm = geometric mean
mn = mean
md = median

habitat types:
f = freshwater
m = marine
t = terrestrial
n = non

taxa groups:
pl = Plants
al = Algae
fu = Fungi
ma = Makro-Invertebrates (", paste0(invertebrates_makro, collapse = ' + '),")
mi = Mikro-Invertebrates (", paste0(invertebrates_mikro, collapse = ' + '),")
da = Daphniidae (Fam.)
cr = Crustacea
ch = Chrionomidae (Fam.)
in = Insects
mo = Molluscs
fi = Fish
ar = Araneae
au = autotroph organisms (", paste0(autotrophs, collapse = ' + '),")
he = heterotroph organisms

You can look up the reference number, casnr, etc. here:
https://cfpub.epa.gov/ecotox/quick_query.htm

---
All values are in ug/L.
---

Note:
In order to build this table, ecological test data on ", su, " substances and ", sp, " species (from ", fa, " families) were processed.\nThe test data were mainly obtained through the EPA ECOTOX data base and some scientific articles and the PPDB.\nIn order to properly classify taxonomic ranks and habitat preferences, the ITIS, NCBI, COL, EOL, NBN and WORMS data bases were queried.\nFor ", ha, " families, additional habitat information (freshwater, marine, terretrial) was collected by hand. No guarantee for perfectness.\nEspecially habitat information can still have wrong classifications.

---
Uploaded: ", Sys.Date(), "';"))

## Maintainance
dbSendQuery(con, paste0("VACUUM ANALYZE  ", schema, ".", table, ";"))

dbDisconnect(con)
dbUnloadDriver(drv)

message('Written to DB:'); freshw_grp
```

```{r cleaning}
rm(su, sp, fa, ha)
```

# Problems
```{r problems. eval=FALSE}
ppdb_epoxi = '135319732' # Kein EPA-Eintrag
ppdb_epoxi_cas = '135319-73-2' # Kein EPA-Eintrag
ppdb_epoxi_cas_old = '106325-08-0'
epa_epoxi = '133855988'

tests[ casnr %in% c(ppdb_epoxi, epa_epoxi) ]
ppdb_l[[ppdb_epoxi_cas]]$etox
```

# Misc
```{r misc, eval=FALSE}
daphnia = tests[ latin_BIname %like% '(?i)Daphnia', .N, latin_BIname ][order(-N)]
fwrite(daphnia, file.path(tempdir(), 'daphnia.csv'))
```