### 2018-07-09 src/EC50aggregation_fin.Rmd

---
title: "Untitled"
author: "AS"
date: "March 29, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

This script creates the table phch_ec50_fin on the basis of phch_ec50 and loads it into the data base.

# Setup
```{r setup}
# packages
require(RPostgreSQL)
require(data.table)
require(stringr)

# variables
cachedir = '/home/andreas/Documents/Projects/species-sensitivity/cache'
plotdir = '/home/andreas/Documents/Projects/species-sensitivity'

# source
source('/home/andreas/Documents/UBA/Project/R/src/DB_access.R')
```

# Load ec50-table
```{r load}
drv = dbDriver("PostgreSQL")
con = dbConnect(drv, user = DBuser, dbname = DBname_1, host = DBhost, port = DBport, password = DBpassword)

ec50 = dbGetQuery(con, "SELECT * FROM phch.phch_ec50")
setDT(ec50)

dbDisconnect(con)
dbUnloadDriver(drv)
```

# Final EC50 columns

## Hierachical partition

### Daphniidae
```{r daphniidae}
ec50_da48_fin = 
  ec50[ ,
        .(casnr = casnr,
          ec50f_da48_fin = ifelse(!is.na(mj50f_da48_vl), mj50f_da48_vl,
                           ifelse(!is.na(pp50f_da48_vl), pp50f_da48_vl,
                           ifelse(!is.na(ep50f_da48_md), ep50f_da48_md, NA))),
          ec50f_da48_tax = ifelse(!is.na(mj50f_da48_vl), mj50f_da48_tax,
                           ifelse(!is.na(pp50f_da48_vl), pp50f_da48_tax,
                           ifelse(!is.na(ep50f_da48_md), ep50f_da48_tax, NA))),
          ec50f_da48_n = ifelse(!is.na(mj50f_da48_vl), mj50f_da48_n,
                         ifelse(!is.na(pp50f_da48_vl), pp50f_da48_n,
                         ifelse(!is.na(ep50f_da48_md), ep50f_da48_n, NA))),
          ec50f_da48_src = ifelse(!is.na(mj50f_da48_vl), 'malaj',
                           ifelse(!is.na(pp50f_da48_vl), 'ppdb',
                           ifelse(!is.na(ep50f_da48_md), 'epa', NA))))]

# keep only the first part of the string (until '-'): This is the minimum value
ec50_da48_fin[ , ec50f_da48_tax := sapply(strsplit(ec50f_da48_tax, '_'), '[', 1)]
```

## Minimum partition

### Algae
```{r algae}
ec50_al_m = 
  melt(ec50, id.vars = 'casnr',
       measure.vars = list(c('mj50f_al4896_vl', 'pp50f_al72_vl', 'ep50f_al4896_min'),
                           c('mj50f_al4896_tax', 'pp50f_al72_tax', 'ep50f_al4896_tax'),
                           c('mj50f_al4896_n', 'pp50f_al72_n', 'ep50f_al4896_n')))

ec50_al_m[ , value4 := ifelse(variable == '1', 'malaj',
                       ifelse(variable == '2', 'ppdb',
                       ifelse(variable == '3', 'epa', NA))) ] #! careful with the order

ec50_al4896_fin = 
  ec50_al_m[ ,
             .(min(value1, na.rm = TRUE),
               .SD[which.min(value1)]$value2,
               .SD[which.min(value1)]$value3,
               .SD[which.min(value1)]$value4),
             casnr  ]

setnames(ec50_al4896_fin, c('casnr', 'ec50f_al4896_fin', 'ec50f_al4896_tax', 'ec50f_al4896_n', 'ec50f_al4896_src'))

# keep only the first part of the string (until '-'): This is the minimum value
ec50_al4896_fin[ , ec50f_al4896_tax := sapply(strsplit(ec50f_al4896_tax, ' - '), '[', 1)]
#ec50_al4896_fin[ , ec50f_al4896_dur := gsub('[^0-9]', '', ec50_al4896_fin$ec50f_al4896_tax) ]
ec50_al4896_fin[ , ec50f_al4896_tax := sapply(strsplit(ec50f_al4896_tax, '_'), '[', 1)]
ec50_al4896_fin[ , ec50f_al4896_tax := sapply(strsplit(ec50f_al4896_tax, '_'), '[', 1)]
ec50_al4896_fin[ , ec50f_al4896_tax := ifelse(str_count(ec50f_al4896_tax, '\\S+') == 1,
                                              paste0(ec50f_al4896_tax, ' sp.'),
                                              ec50f_al4896_tax)]

for (i in names(ec50_al4896_fin)) {
  ec50_al4896_fin[ get(i) == Inf, (i) := NA ]
  ec50_al4896_fin[ get(i) == '', (i) := NA ]
}
rm(ec50_al_m)
```

### Invertebrates
```{r makro-invertebrates}
ec50_ma_m = 
  melt(ec50, id.vars = 'casnr',
       measure.vars = list(c('mj50f_da48_vl', 'pp50f_iv4896_min', 'ep50f_ma4896_min'),
                           c('mj50f_da48_tax', 'pp50f_iv4896_tax', 'ep50f_ma4896_tax'),
                           c('mj50f_da48_n', 'pp50f_iv4896_n', 'ep50f_ma4896_n')))

ec50_ma_m[ , value4 := ifelse(variable == '1', 'malaj',
                       ifelse(variable == '2', 'ppdb',
                       ifelse(variable == '3', 'epa', NA))) ] #! careful with the order

ec50_ma4896_fin = 
  ec50_ma_m[ ,
             .(min(value1, na.rm = TRUE),
               .SD[which.min(value1)]$value2,
               .SD[which.min(value1)]$value3,
               .SD[which.min(value1)]$value4),
             casnr  ]

setnames(ec50_ma4896_fin, c('casnr', 'ec50f_ma4896_fin', 'ec50f_ma4896_tax', 'ec50f_ma4896_n', 'ec50f_ma4896_src'))

# keep only the first part of the string (until '-'): This is the minimum value
ec50_ma4896_fin[ , ec50f_ma4896_tax := sapply(strsplit(ec50f_ma4896_tax, ' - '), '[', 1)]
#ec50_ma4896_fin[ , ec50f_ma4896_dur := gsub('[^0-9]', '', ec50_ma4896_fin$ec50f_ma4896_tax) ]
ec50_ma4896_fin[ , ec50f_ma4896_tax := sapply(strsplit(ec50f_ma4896_tax, '_'), '[', 1)]
ec50_ma4896_fin[ , ec50f_ma4896_tax := ifelse(str_count(ec50f_ma4896_tax, '\\S+') == 1,
                                              paste0(ec50f_ma4896_tax, ' sp.'),
                                              ec50f_ma4896_tax)]

for (i in names(ec50_ma4896_fin)) {
  ec50_ma4896_fin[ get(i) == Inf, (i) := NA ]
  ec50_ma4896_fin[ get(i) == '', (i) := NA ]
}
rm(ec50_ma_m)
```

### Fish
```{r fish}
ec50_fi_m = 
  melt(ec50, id.vars = 'casnr',
       measure.vars = list(c('mj50f_fi96_vl', 'pp50f_fi96_vl', 'ep50f_fi96_min'),
                           c('mj50f_fi96_tax', 'pp50f_fi96_tax', 'ep50f_fi96_tax'),
                           c('mj50f_fi96_n', 'pp50f_fi96_n', 'ep50f_fi96_n')))

ec50_fi_m[ , value4 := ifelse(variable == '1', 'malaj',
                       ifelse(variable == '2', 'ppdb',
                       ifelse(variable == '3', 'epa', NA))) ] #! careful with the order

ec50_fi96_fin = 
  ec50_fi_m[ ,
             .(min(value1, na.rm = TRUE),
               .SD[which.min(value1)]$value2,
               .SD[which.min(value1)]$value3,
               .SD[which.min(value1)]$value4),
             casnr  ]

setnames(ec50_fi96_fin, c('casnr', 'ec50f_fi96_fin', 'ec50f_fi96_tax', 'ec50f_fi96_n', 'ec50f_fi96_src'))

# keep only the first part of the string (until '-'): This is the minimum value
ec50_fi96_fin[ , ec50f_fi96_tax := sapply(strsplit(ec50f_fi96_tax, ' - '), '[', 1)]
#ec50_fi96_fin[ , ec50f_fi96_dur := gsub('[^0-9]', '', ec50_fi96_fin$ec50f_fi96_tax) ]
ec50_fi96_fin[ , ec50f_fi96_tax := sapply(strsplit(ec50f_fi96_tax, '_'), '[', 1)]
ec50_fi96_fin[ , ec50f_fi96_tax := ifelse(str_count(ec50f_fi96_tax, '\\S+') == 1,
                                              paste0(ec50f_fi96_tax, ' sp.'),
                                              ec50f_fi96_tax)]

for (i in names(ec50_fi96_fin)) {
  ec50_fi96_fin[ get(i) == Inf, (i) := NA ]
  ec50_fi96_fin[ get(i) == '', (i) := NA ]
}
rm(ec50_fi_m)
```

## Merge 1
```{r merge1}
fin_dts = grep('_fin', ls(), value = TRUE, ignore.case = TRUE)
fin_l = lapply(fin_dts, get)
names(fin_l) = fin_dts

ec50f_grp = Reduce(function(...) merge(..., on = 'casnr', all = TRUE), fin_l)
ec50f_grp[ec50, on = 'casnr', subst_name := i.subst_name]
setcolorder(ec50f_grp, c(1, ncol(ec50f_grp), 3:ncol(ec50f_grp)-1))
```

# Write to data base

```{r write-ec50f_fin}
drv = dbDriver("PostgreSQL")
con = dbConnect(drv, user = DBuser, dbname = DBname_1, host = DBhost, port = DBport, password = DBpassword)

schema = 'phch'
table = 'phch_ec50_fin'

## upload
dbSendQuery(con, paste0("DROP TABLE IF EXISTS  ", schema, ".", table, ";"))
dbWriteTable(con, ec50f_grp,
             name = c(schema, table), row.names = FALSE)

## User rights
dbSendQuery(con, paste0("GRANT ALL ON TABLE ", schema, ".", table, " TO scharmueller;
                  GRANT ALL ON TABLE ", schema, ".", table, " TO bfg;
                  GRANT SELECT ON TABLE ", schema, ".", table, " TO bfg_read;"))

## Primary Key
# dbSendQuery(con, paste0("ALTER TABLE ", schema, ".", table, " ADD PRIMARY KEY (variable_id);"))

## COMMENT ON
dbSendQuery(con, paste0("COMMENT ON TABLE  ", schema, ".", table, "  IS '
final ec50 table with aggregated Invertebrate values based on phch.phch_ec50

(1) ec50f_al4896_fin = minimum from mj50f_al4896_vl, pp50f_al72_vl, ep50f_al4896_min
(2) ec50f_da48_fin = if present: mj50f_al4896_vl, pp50f_al72_vl, ep50f_al4896_md
(3) ec50f_fi96_fin = minimum from mj50f_da48_vl, pp50f_ma48_min, ep50f_ma4896_min
(4) ec50f_ma4896_fin = minimum from mj50f_da48_vl, pp50f_ma48_min, ep50f_ma4896_min

---
Abbreviations:
suffixes:
_fin = final aggregated value
_tax = Taxon corresponding to _fin value
_n = number of vlaues/tests that were used in the aggregation
_src = source (malaj, ppdb or epa)

taxa groups:
pl = Plants
al = Algae
fu = Fungi
ma = Makro-Invertebrates (Annelida + Echinodermata + Mollusca + Nemertea + Platyhelminthes + Porifera + Crustacea + Arachnida + Diplopoda + Entognatha + Insecta)
mi = Mikro-Invertebrates (Bryozoa + Chaetognatha + Ciliophora + Cnidaria + Gastrotricha + Nematoda + Rotifera)
da = Daphniidae (Fam.)
cr = Crustacea
ch = Chrionomidae (Fam.)
in = Insects
mo = Molluscs
fi = Fish
ar = Araneae
au = autotroph organisms (Plants + Algae + Bryophyta)
he = heterotroph organisms

---
All values are in ug/L.

Uploaded: ", Sys.Date(), "';"))

## Maintainance
dbSendQuery(con, paste0("VACUUM ANALYZE  ", schema, ".", table, ";"))

dbDisconnect(con)
dbUnloadDriver(drv)
```