### 2018-07-09 src/EC50comparison_plots.Rmd

---
title: "Untitled"
author: "AS"
date: "April 2, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup}
# packages
require(data.table)
require(RPostgreSQL)
require(ggplot2)

# variables
cachedir = '/home/andreas/Documents/Projects/species-sensitivity/cache'
plotdir = '/home/andreas/Documents/Projects/species-sensitivity/plots'

# source
source('/home/andreas/Documents/UBA/Project/R/src/DB_access.R')
```

# Load test results
```{r load-tests}
tests3 = readRDS(file.path(cachedir, 'tests3.rds'))
psm = readRDS(file.path(cachedir, 'psm2.rds'))
```

```{r load-ec50s}
drv = dbDriver("PostgreSQL")
con = dbConnect(drv, user = DBuser, dbname = DBname_1, host = DBhost, port = DBport, password = DBpassword)

ec50 = dbGetQuery(con, "SELECT * FROM phch.phch_ec50")
setDT(ec50)

dbDisconnect(con)
dbUnloadDriver(drv)

ec50[psm, on = 'casnr', psm_type := i.psm_type]
ec50 = ec50[psm_type %in% c('fungicide', 'herbicide', 'insecticide')]
```

# Test durations per organism groups
```{r duration-taxon}
tests_epa = tests3[ source == 'epa_ecotox' & isFre_fin == '1' ]

# merging taxa that are lower than supgroup2:
dt_ch = tests_epa[ family == 'Chironomidae' , .N, duration2][, grp := 'Chironomidae']
setnames(dt_ch, c('dur', 'N', 'grp'))
dt_da = tests_epa[ family == 'Daphniidae' , .N, duration2][, grp := 'Daphniidae']
setnames(dt_da, c('dur', 'N', 'grp'))
dt_cr = tests_epa[ supgroup == 'Crustacea' , .N, duration2][, grp := 'Crustacea']
setnames(dt_cr, c('dur', 'N', 'grp'))
dt_mo = tests_epa[ supgroup == 'Mollusca' , .N, duration2][, grp := 'Mollusca']
setnames(dt_mo, c('dur', 'N', 'grp'))
dt_in = tests_epa[ supgroup == 'Insecta' , .N, duration2][, grp := 'Insecta']
setnames(dt_in, c('dur', 'N', 'grp'))

dt_dur = tests_epa[ , .N, .(supgroup2, duration2)]
setnames(dt_dur, c('grp', 'dur', 'N'))

dt_dur = rbindlist(list(dt_dur, dt_ch, dt_da, dt_cr, dt_mo, dt_in), use.names = TRUE)

# plot
gg_dur = ggplot(dt_dur[ !grp %in% c('Aves', 'Mammalia', NA)], aes(y = N, x = dur)) +
  geom_label(aes(label = dur)) +
  facet_wrap( ~ grp, scales = 'free') +
  labs(title = 'Duration of tests per taxon',
       subtitle = 'freshwater organisms') +
  theme_bw()

ggsave(gg_dur, filename = file.path(plotdir, 'duration-taxon.png'),
       width = 10, height = 7)

# Why are there NAs???!!
# 'cause it's not yet classified in the lookup table
tests3[ source == 'epa_ecotox' & is.na(supgroup2)][ , .N, latin_BIname]
```

# Effect endpoint per taxon
```{r effect-taxon}
dt_eff = tests_epa[ , .N, .(supgroup2, effect)]
setnames(dt_eff, c('grp', 'effect', 'N'))

gg_eff = ggplot(dt_eff[ !grp %in% c('Mammalia', 'Aves', NA) ], aes(y = N, x = effect)) +
  geom_point() +
  facet_wrap( ~ grp, scales = 'free') +
  labs(title = 'Effect numbers per taxon',
       subtitle = 'freshwater organisms') +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1.1))

ggsave(gg_eff, filename = file.path(plotdir, 'effect-taxon.png'),
       width = 10, height = 7)
```

# Compare 24-120 tests to specific times
```{r comparison-plots2}
## Algae
# data
al_cols = grep('al.+(vl|gm)', names(ec50), value = TRUE)

ec50al = ec50[ , .SD, .SDcols = c('casnr', 'psm_type', al_cols)]
ec50al = ec50al[ psm_type %in% c('fungicide', 'herbicide', 'insecticide')]
ec50al_m = melt(ec50al, id.var = c('casnr', 'psm_type')) # melt
ec50al_m72 = melt(ec50al[ ,!c('psm_type')], id.vars = c('casnr', 'pp50f_al72_vl')) # range per duratiion group

# plot 1 [doesn't help much]
ggplot(ec50al_m, aes(y = reorder(casnr, value), x = value, col = variable)) +
  geom_point() +
  scale_x_log10() +
  facet_wrap( ~ variable, scales = 'free_x', ncol = 3)

# plpt 2 [doesn't help much]
ec50al_m_rng = ec50al_m[ , .(rng = range(value, na.rm = TRUE)[2] -
                               range(value, na.rm = TRUE)[1]),
                         casnr]
ec50al_m_rng[ is.infinite(rng), rng := NA ]

ggplot(ec50al_m_rng, aes(y = reorder(casnr, rng), x = rng)) +
  geom_point() +
  coord_cartesian(xlim = c(0,1000))
  #scale_x_log10()

# plot 3 [doesn't help much]
# sum per psm_type and variable
ec50al_m_sum = ec50al_m[ , .(mn = mean(value, na.rm = TRUE),
                             sd = sd(value, na.rm = TRUE)), .(variable, psm_type) ]

ggplot(ec50al_m_sum, aes(y = mn, variable)) +
  geom_point() +
  geom_errorbar(aes(ymin = mn-sd, ymax = mn+sd)) +
  facet_wrap( ~ psm_type) +
  coord_flip() +
  theme_bw()

# plot 4 [doesn't help much]
cols = grep('casnr|al.+(vl|gm)', names(ec50), value = TRUE)#, 'pp50f_al72_vl')
ec50al = ec50[ , .SD, .SDcols = cols]
cols = setdiff(cols, c('pp50f_al72_vl', 'casnr'))
ec50al_diff = copy(ec50al)[ , (cols) := abs(pp50f_al72_vl - .SD), .SDcols = cols]

ec50al_diff_m = melt(ec50al_diff, id.vars = 'casnr')

ggplot(ec50al_diff_m, aes(y = casnr, x = value)) +
  geom_point() +
  facet_wrap( ~ variable)

```