### 2018-07-09 src/source_malaj.Rmd

---
title: "Source Malaj data"
author: "Andreas Scharmüller"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
editor_options:
  chunk_output_type: console
---

# Setup
```{r setup}
# packages
require(data.table)

# variables
cachedir = '/home/andreas/Documents/Projects/species-sensitivity/cache'
fundir = '/home/andreas/Documents/Projects/species-sensitivity/functions'

# source
source(file.path(fundir, 'casconv.R'))
```

# Load PSM data
```{r load-psm}
psm = readRDS(file.path(cachedir, 'psm2.rds'))
```

# Load Malaj data
```{r load-malaj}
malaj = fread('/home/andreas/Documents/data/malaj2014/Toxicity_data_malaj_2014.csv')
```

## Data cleaning
```{r malaj-cleaning}
names(malaj) = tolower(names(malaj))
malaj = malaj[ ,c('casnumber', 'cas', 'lc50 dm', 'source details_dm', 'type of data_dm', 'lc50 pim', 'source details_pim', 'type of data_pim', 'lc50 sele', 'source details_sele', 'type of data_sele', 'code', 'determinand_hazsubs')]
setnames(malaj, c('casnr', 'cas', 'lc50_dm', 'lc50_dm_source', 'lc50_dm_type', 'lc50_pim', 'lc50_pim_source', 'lc50_pim_type', 'lc50_sel', 'lc50_sel_source', 'lc50_sel_type', 'psm_type', 'subst_name'))

# cas
malaj = malaj[ !is.na(casnr) ] # no single pesticides
malaj[ , cas := casconv(casnr) ]

# replace '' with NA
for(i in names(malaj)) {
  malaj[get(i) == '', (i) := NA ]
}

# taxa
malaj[ , `:=` (dm = 'Daphnia magna',
               pp = 'Pimephales promelas',
               ps = 'Pseudokirchneriella subcapitata')]

# melt
malaj_m = melt(malaj,
               id.vars = c('casnr', 'cas', 'psm_type', 'subst_name'),
               measure.vars = list(
                 c('dm', 'pp', 'ps'),
                 c('lc50_dm', 'lc50_pim', 'lc50_sel'),
                 c('lc50_dm_source', 'lc50_pim_source', 'lc50_sel_source'),
                 c('lc50_dm_type', 'lc50_pim_type', 'lc50_sel_type')),
               value.name = c('latin_BIname', 'value', 'source', 'type'))

# other variables
malaj_m[ , `:=` (endpoint = 'lc50',
                 variable_id = NA,
                 unit = 'ug/L',
                 duration = ifelse(latin_BIname == 'Daphnia magna', 48,
                            ifelse(latin_BIname == 'Pimephales promelas', 96,
                            ifelse(latin_BIname == 'Pseudokirchneriella subcapitata', 4896, NA))),
                 source = 'malaj')]
malaj_m[ , variable := NULL]

malaj_m = malaj_m[ !is.na(value) ]

# column order
setcolorder(malaj_m, c('variable_id', 'subst_name', 'casnr', 'cas', 'value', 'unit', 'psm_type', 'source', 'type', 'duration', 'endpoint', 'latin_BIname'))
```

## Filter
```{r filter}
malaj_m = malaj_m[ !type %in% c('B', 'P/1', 'P/2') ] 
```

# Cleaning
```{r cleaning}
rm(malaj, psm)
```