### 2018-07-09 R/EC50aggregation_epa2.Rmd
renamed folder <src> to <R>

---
title: "Untitled"
author: "AS"
date: "April 5, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

This script is loaded by EC50aggregation.Rmd and aggregates the EPA data. There is also EC50aggregation_epa.Rmd which does the aggregation differently, by excluding outliers and not taking the median.

# Setup
```{r setup}
# packages
require(data.table)

# cachedir
cachedir = '/home/andreas/Documents/Projects/species-sensitivity/cache'
fundir = '/home/andreas/Documents/Projects/species-sensitivity/functions'
plotdir = '/home/andreas/Documents/Projects/species-sensitivity/plots'

# functions
source(file.path(fundir, 'agg_group2.R'))
```

# Load
```{r load}
tests3 = readRDS(file.path(cachedir, 'tests3.rds'))
```

# Filter
```{r epa}
epa3 = tests3[ source == 'epa_ecotox' ]
epa3 = epa3[ !supgroup %in% c('Entognatha', 'Arachnida') ]
```

# Aggregate

## Freshwater organism groups
### Input data
```{r AGGREGATE-epa-fresh-todo}
## create variable parameter list
tol_f =
  list(grp_names = c('Chironomidae96', 'Daphnia48', 'Insecta4896', 'Crustacea4896', 'Annelida4896', 'Platyhelminthes4896', 'Mollusca4896', 'Makro_Inv4896', 'Fish96', 'Algae4896', 'Plants168'),

       grp_par = c(rep('family', 2), rep('supgroup', 5), rep('supgroup2', 4)),

       grp_arg = c('Chironomidae', 'Daphniidae', 'Insecta', 'Crustacea', 'Annelida', 'Platyhelminthes', 'Mollusca', 'Makro_Inv', 'Fish', 'Algae', 'Plants'),

       dur_par = rep('duration', 11),

       dur_arg = list(96, 48, c(48,96), c(48,96), c(48,96), c(48,96), c(48,96), c(48,96), 96, c(48,96), 168))

freshw_grp = as.data.table(tol_f); freshw_grp
```

##### Output data
```{r AGGREGATE-epa-fresh-mapply-data}
ep50f_l =
  mapply(agg_group2,
         grp_par = tol_f$grp_par, grp_arg = tol_f$grp_arg, dur_arg = tol_f$dur_arg, # variable arguments
         MoreArgs = list(dt = epa3, habi = 'fresh', dur_par = 'duration'), # fixed arguments
         SIMPLIFY = FALSE)
names(ep50f_l) = tol_f$grp_names

ep50f_agg = Reduce(function(...) merge(..., by = 'casnr', all = TRUE), ep50f_l)
ep50f_agg[ ep50f_agg == Inf ] = NA; ep50f_agg[ ep50f_agg == NaN ] = NA;
fwrite(ep50f_agg, file.path(tempdir(), 'ep50f_agg.csv')) # debuging
```