### 2018-07-09 R/speciesinfo.Rmd

---
title: "speciesinfo"
author: "AS"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Setup
```{r setup}
## packages
require(data.table)
require(taxize)

online = F

cachedir = '/home/andreas/Documents/Projects/species-sensitivity/cache'
```

# TODO
```{r TODO}
tests_todo = readRDS(file.path(cachedir, 'tests1.rds'))
todo_taxa = sort(unique(tests_todo$latin_BIname))
rm(tests_todo)
```

# QUERIES

## ITIS, NCBI, COL

As the EPA ECOTOX organism groups (algae, plants, crustaceans, fish, insects, invertebrates and molluscs) are not concise we establish new groups querying taxa information using taxize.

```{r QUERY-ITIS-COL-NBN-taxa-groups, eval=TRUE, echo=TRUE}
time_tot = Sys.time()
#### Query ITIS ----
if (online) {
  itis_todo = todo_taxa
  # itis_todo = itis_todo[1:5] # debug me!

  time = Sys.time()
  result_itis = classification(itis_todo, db = 'itis', rows = 1)
  message('Query took: ', format(Sys.time() - time, format = '%M')) # 8 min

  saveRDS(result_itis, file.path(cachedir, 'ITIS_taxa_result.rds'))

} else {
  result_itis = readRDS(file.path(cachedir, 'ITIS_taxa_result.rds'))
}

# add itis identifier (to distinguish later between ITIS and NCBI)
result_itis[!is.na(result_itis)] = mapply(cbind, result_itis[!is.na(result_itis)], source = 'itis', SIMPLIFY = FALSE) # don't spoil the NAs

# Entries that are not found in ITIS
todo2_1 = sort(names(result_itis[ is.na(result_itis) ]))

# Matches that don't have a family entry
todo2_2 = names(which(unlist(lapply(
  lapply(result_itis[ !is.na(result_itis) ],
         function(x) x[ x$rank == 'family', ]$name),
  length)) == 0))

# Only keep valid results
result_itis = result_itis[ !is.na(result_itis) & !names(result_itis) %in% todo2_2 ]

# todo NCBI
todo2 = c(todo2_1, todo2_2)

#### Query COL (Catalogue of Life) - seems to be the fastest ----
if (online) {
  time = Sys.time()
  result_col = classification(todo2, db = 'col', rows = 1)  
  message('Query took: ', format(Sys.time() - time, format = '%M')) # 6 sec

  saveRDS(result_col, file.path(cachedir, 'COL_taxa_result_NOT_IN_ITIS.rds'))

} else {
  result_col = readRDS(file.path(cachedir, 'COL_taxa_result_NOT_IN_ITIS.rds'))
}

# add identifier col
result_col[!is.na(result_col)] = lapply(result_col[!is.na(result_col)], cbind, source = 'col')

# Entries that are not found
todo3_1 = sort(names(result_col[ is.na(result_col) ]))

# Matchin entries that don't have a family entry
todo3_2 = names(which(unlist(lapply(
  lapply(result_col[ !is.na(result_col) ],
         function(x) x[ x$rank == 'family', ]$name),
  length)) == 0))

todo3 = c(todo3_1, todo3_2)

# Only keep valid results
result_col = result_col[ !is.na(result_col) & !names(result_col) %in% todo3_2 ]

#### Query NBN (National Biodiversity Network UK) ----
if (online) {
  time = Sys.time()
  result_nbn = classification(todo3, db = 'nbn', rows = 1)
  message('Query took: ', format(Sys.time() - time, format = '%M'))

  saveRDS(result_nbn, file.path(cachedir, 'NBN_taxa_result_NOT_IN_ITIS_COL.rds'))

} else {
  result_nbn = readRDS(file.path(cachedir, 'NBN_taxa_result_NOT_IN_ITIS_COL.rds'))
}

# add ifentifier nbn
result_nbn[ !is.na(result_nbn) ] = lapply(result_nbn[ !is.na(result_nbn) ], cbind, source = 'nbn')

# Entries that are not found
todo4_1 = sort(names(result_nbn[ is.na(result_nbn) ]))

# Matchin entries that don't have a family entry
todo4_2 = names(which(unlist(lapply(
  lapply(result_nbn[ !is.na(result_nbn) ],
         function(x) x[ x$rank == 'family', ]$name),
  length)) == 0))

todo4 = c(todo4_1, todo4_2)

# Only keep valid results
result_nbn = result_nbn[ !is.na(result_nbn) & !names(result_nbn) %in% todo4_2 ]

#### TOL (Open Tree of Life) ----
if (online) {
  time = Sys.time()
  result_tol = classification(todo4, db = 'tol', rows = 1)
  message('Query took: ', format(Sys.time() - time, format = '%M')) # 2.8 min

  saveRDS(result_tol, file.path(cachedir, 'TOL_taxa_result_NOT_IN_ITIS_COL_NBN.rds'))

} else {
  result_tol = readRDS(file.path(cachedir, 'TOL_taxa_result_NOT_IN_ITIS_COL_NBN.rds'))
}

# add identifier: ncbi
result_tol[!is.na(result_tol)] = mapply(cbind, result_tol[!is.na(result_tol)], source = 'tol', SIMPLIFY = FALSE)

# Entries that are not found
todo5_1 = sort(names(result_tol[ is.na(result_tol) ]))

# Matches that don't have a family entry
todo5_2 = names(which(unlist(lapply(
  lapply(result_tol[ !is.na(result_tol) ],
         function(x) x[ x$rank == 'family', ]$name),
  length)) == 0))

todo5 = c(todo5_1, todo5_2)

# Only keep valid results
result_tol = result_tol[ !is.na(result_tol) & !names(result_tol) %in% todo5 ]

message('The queries took: ', format(Sys.time() - time_tot, format = '%M'))
```

## not found & errata
```{r QUERY-not-found}
# sort(todo5)
# Taxa that couldn't be identified with the above methods
##DEPRECATED!!!!
# leftovers = list(`Agamermis unka` = data.table(name = 'Mermithidae',
#                                                         rank = 'family',
#                                                         id = NA,
#                                                         source = 'by_hand'),
#                  Neoniphargus = data.table(name = 'Neoniphargidae',
#                                                      rank = 'family',
#                                                      id = NA,
#                                                      source = 'by_hand'),
#                  Newnhamia = data.table(name = 'Notodromadidae',
#                                         rank = 'family',
#                                         id = NA,
#                                         source = 'by_hand'),
#                  `Parachlorella kessleri` = data.table(name = 'Chlorellaceae',
#                                            rank = 'family',
#                                            id = NA,
#                                            source = 'by_hand'),
#                  `Paramelita nigroculus` = data.table(name = 'Paramelitidae',
#                                                         rank = 'family',
#                                                         id = NA,
#                                                         source = 'by_hand'),
#                  Sphaerodema = data.table(name = 'Belostomatidae',
#                                           rank = 'family',
#                                           id = NA,
#                                           source = 'by_hand'),
leftovers = list(`Supputius cincticeps` = data.table(name = 'Pentatomoidea',
                                           rank = 'family',
                                           id = NA,
                                           source = 'by_hand'),
                 `Tetrahymena thermophila` = data.table(name = 'Tetrahymenidae',
                                                        rank = 'family',
                                                        id = NA,
                                                        source = 'by_hand'))
```

# Bind
```{r bind}
result_taxize = c(result_itis,
                  result_col,
                  result_nbn,
                  result_tol,
                  leftovers)

## Retreive name, family, source
# NOT very fast, NOT very R, but not as clumsy as lapply ;)
tx_list = list()
for (i in 1:length(result_taxize)) {
  dt = as.data.table(result_taxize[[i]], stringsAsFactors = FALSE)
  dt[ , source := as.character(source) ]
  latin_BIname = names(result_taxize[i])

  if (nrow(dt[ rank == 'family' ]) == 0) {
    tx = data.table(latin_BIname = latin_BIname,
                    family = 'NOT',
                    source_tax = unlist(dt[ nrow(dt), 'source' ]),
                    stringsAsFactors = FALSE)
  } else {
    tx = data.table(latin_BIname = latin_BIname,
                    family = unlist(dt[ rank == 'family', 'name' ]),
                    source_tax = unlist(dt[ nrow(dt), 'source' ]),
                    stringsAsFactors = FALSE)
  }
  tx_list[[i]] = tx
}

tx_dt = rbindlist(tx_list)
setnames(tx_dt, c("latin_BIname", "family_tax", "source_tax"))

# checks
family_check =
  tx_dt[ is.na(family_tax) | family_tax == '' ]

if (nrow(family_check) != 0) {
  warning(nrow(family_check), ' missing family names.')
}
```

# Errata

Final errata mostly due to change in taxonomic hierachies and names.
```{r errata}
# COL: 'not assigned'
tx_dt[ latin_BIname == 'Ceriodaphnia silvestrii', `:=`
       (family_tax = 'Daphniidae',
        source_tax = 'by_hand')]
tx_dt[ latin_BIname == 'Daphnia spinulata', `:=`
       (family_tax = 'Daphniidae',
        source_tax = 'by_hand')]
tx_dt[ latin_BIname == 'Oculimacula yallundae', `:=`
       (family_tax = 'Dermateaceae',
        source_tax = 'by_hand')]
tx_dt[ latin_BIname == 'Pseudosida ramosa', `:=`
       (family_tax = 'Sididae',
        source_tax = 'by_hand')] 	
tx_dt[ latin_BIname == 'Simocephalus elizabethae', `:=`
       (family_tax = 'Daphniidae',
        source_tax = 'by_hand')]

# Wrongly classified by taxize (from PPDB or Malaj)
tx_dt[ latin_BIname == 'Navicula seminulum', `:=`
       (family_tax = 'Naviculaceae',
        source_tax = 'by_hand')]

# Other:
tx_dt[ family_tax == 'Melolonthidae', `:=`
       (family_tax = 'Scarabaeidae',
        source_tax = 'by_hand') ]
tx_dt[ family_tax == 'Megaperidae', `:=`
       (family_tax = 'Apocreadiidae',
        source_tax = 'by_hand') ]
tx_dt[ family_tax == 'Pentatomoidea', `:=` # superfamily_tax
       (family_tax = 'Pentatomidae',
        source_tax = 'by_hand') ]
tx_dt[ family_tax == 'Lymantriidae', `:=` # subfamily_tax
       (family_tax = 'Erebidae',
        source_tax = 'by_hand')]
tx_dt[ family_tax == 'Filiniidae', `:=` # unaccepted synonym
       (family_tax = 'Trochosphaeridae',
        source_tax = 'by_hand')]
```

# Cleaning
```{r cleaning}
oldw = getOption("warn")
options(warn = -1)

rm(dt, family_check, tx_list, tx,
   leftovers, result_itis, result_col, result_nbn, result_tol,
   list = grep('todo', ls(), value = TRUE),
   latin_BIname, online, time_tot, i)

options(warn = oldw)
```