### 2018-08-02 TeX/organigram2.tex
added .tex for organigram of data preparation

\documentclass{article}
%\usepackage{graphicx} % includegraphics
\usepackage{tikz}
\usetikzlibrary{trees, arrows, shapes.geometric} % TikZ libraries
\usepackage{lscape}

\begin{document}

\thispagestyle{empty} % removes the 1

% tree styles (defined by template)
\tikzstyle{every node}=[draw=black,thick,anchor=west]
\tikzstyle{selected}=[draw=red,fill=red!30]
\tikzstyle{optional}=[dashed,fill=gray!50]
% self defined
\tikzstyle{io}=[rectangle, rounded corners, minimum width = 2cm, minimum height = 1cm, text centered, draw = black, fill = blue!10]
\tikzstyle{main}=[rectangle, rounded corners, minimum width = 3cm, minimum height = 1cm, text centered, draw = black, fill = red!30]
\tikzstyle{ids}=[diamond, minimum width = 3cm, minimum height = 1cm, text centered, draw = black, fill = green!30]
\tikzstyle{res}=[rectangle, rounded corners, minimum width = 2cm, minimum height = 1cm, text centered, draw = black, fill = blue!30]

%arrows
\tikzstyle{arrow} = [thick, ->, > = stealth]

\begin{landscape}
%\parindent=0pt % extra spaces towards the edges
%\eject\null
%\vfill
%\hspace*{-4cm}

\begin{tikzpicture}[
  grow via three points={one child at (0.5,-1) and
  two children at (0.5,-1) and (0.5,-1.7)},
  edge from parent path={(\tikzparentnode.south) |- (\tikzchildnode.west)},
  remember picture, overlay, % r.p. save picture size in the picture, o. puts picture out of text segm.
  inner sep = 0pt, outer sep = 0pt
]

%%% nodes %%%
%% input
% TODO coordinates (-3,4) is strange. Also rectangle around current page seems wrong
\node[text=red] (io) [io, align = left] at (-3,4) {Table of Chemicals \\ CAS, Compound name}; % coordinates: (current page.east)

\draw[thick] (current page.south west) rectangle (current page.north east); % debuging

%% chemicals
% id nodes
\node (cas) [ids, below of=io, yshift = -4cm] {CAS};
% main nodes
\node (epa) [main, right of=cas, yshift=1cm, xshift=3cm] {EPA ectox tests}
	child { node { qu\_epa.R} };
\node (chemscr) [main, right of=cas, yshift=-1cm, xshift=3cm] {Chemical scripts ch}
	child { node { Alan Wood Pesticide Compendium - \textbf{qu\_aw.R } }}
	child { node { Pesticide Action Network - \textbf{ch\_pp.R} }}
	child { node { PhysProp - \textbf{ch\_pp.R} }}
	child { node { PubChem - \textbf{qu\_pubchem.R} }
		child { node { Variable - InchiKey }} };
% result nodes
\node (toxres) [res, right of=cas, yshift=0cm, xshift=7cm] {ETOX Test Results}; % result
%\includegraphics{/home/andreas/Documents/Projects/etox-base/data/pubchemlogo2017.png};

%\animategraphics{https://cfpub.epa.gov/ecotox/epafiles/images/logo_epaseal.gif}

%% organisms
% id nodes
\node (taxon) [ids, below of=cas, yshift=-9cm, align = center] {Taxon}; % \\ {(Genus or Species)}}; % escape brackets usin curly brackets
% main nodes
\node (taxonquery) [main, right of=taxon, xshift=3cm, yshift=0cm] {Taxonomic data};
\node (habitat) [main, right of=taxon, xshift=3cm, yshift=1.5cm] {Habitat data}
	child { node { WORMS data base - \textbf{tx\_worms.R} }
		child { node { Family }}
		child { node { Species }} }
	child [dashed] { node { Freshwaterecology.info data }} ;
\node (region) [main, right of = taxon, xshift=3cm, yshift=-1.5cm] {Regional data};
% result nodes
\node (taxres) [res, right of=taxon, xshift=7cm] {Taxon and Habitat Result}; % result

%% final result
\node (testres) [res] at (15,-5) {Final result};

%%% arrows %%%
\draw [arrow] (io) -- (cas);
\draw [->] (cas) -- node[anchor = east] {receive} (chemscr);
\draw [->, dashed] (cas) -- node[anchor = south] {receive} (epa);
\draw [->] (epa) -- (toxres);
\draw [->] (chemscr) -- (toxres);
\draw [->] (toxres) .. controls (5,-6) and (1,-4) .. (taxon);
\draw [->] (taxon) -- (habitat);
\draw [->] (taxon) -- (region);
\draw [->] (taxon) -- (taxonquery);
\draw [->] (habitat) -- (taxres);
\draw [->] (taxonquery) -- (taxres);
\draw [->] (region) -- (taxres);
\draw [->] (toxres) -- (testres);
\draw [->] (taxres) -- (testres);
\draw [->, dashed] (testres) -- (20,-5);

\end{tikzpicture}
\hfill
\vfill
\end{landscape}

\clearpage

\begin{landscape}
\begin{tikzpicture}

%% entities
\node [res] (clean) at (-5,5) {Cleaning};
\node (filter) [res, right of = clean, xshift = 3cm] {Filter};
\node (agg1) [res, right of = filter, xshift = 3cm] {Aggregation 1};
\node (agg2) [res, right of = agg1, xshift = 3cm] {Aggregation 2};

%% arrows
\draw [->, dashed] (-5,8) -- (clean);

%% boxes
\draw [dashed] (1,6) rectangle (3,-11);
\draw [dashed] (4,5) rectangle (7,-11);
\draw [dashed] (9,5) rectangle (11,-11);
\draw [dashed] (14,5) rectangle (15,-11);

\end{tikzpicture}
\end{landscape}

\end{document}