### 2019-02-05 article/tikz/tree_postgres.tex
Updates from Overleaf

\documentclass[border=5pt]{standalone}

\usepackage{forest}

\begin{document}
\begin{forest}
for tree={font=\ttfamily, grow'=0, child anchor=west, parent anchor=south, anchor=west, calign=first,
    inner xsep=7pt,
    edge path={
      \noexpand\path [draw, \forestoption{edge}]
      (!u.south west) +(7.5pt,0) |- (.child anchor) pic {folder} \forestoption{edge label};
    },
    before typesetting nodes={
      if n=1
        {insert before={[,phantom]}}
        {}
    },
    fit=band,
    before computing xy={l=15pt}
}

[Postgres data base
    [etoxYYYYMMDD
        [ecotox]
        [public]
    ]
]

\end{forest}
\end{document}