### 2019-02-05 article/test.tex
Updates from Overleaf

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{csvsimple}
\usepackage{pgfplotstable}

\begin{document}

\begin{table}
\pgfplotstabletypeset[
	col sep=comma,
	%string type,
	columns={description,script}
	columns/description/.style={column name=Description, column type={|l}},
	columns/script/.style={column name=Script, column type={|l}},
	%columns/name/.style={column name=Name, column type={|l}},
	%columns/surname/.style={column name=Surname, column type={|l}},
	%columns/age/.style={column name=Age, column type={|c|}},
	%every head row/.style={before row=\hline,after row=\hline},
	%every last row/.style={after row=\hline},
	]{data/scripts.csv}

\end{table}

\pgfplotstabletypeset[
    col sep=comma,
    string type,
    columns={Script,Description},
    columns/Description/.style={column type={|l}},
    columns/Script/.style={column type={|l}, % needed for replacing underscore
                           postproc cell content/.code={%
                                \pgfplotsutilstrreplace{_}{\_}{##1}%
                                \pgfkeyslet{/pgfplots/table/@cell content}\pgfplotsretval
                                },
                            },
    every head row/.style={before row=\hline,after row=\hline},
	every last row/.style={after row=\hline},
    ]{data/scripts.csv}

\end{document}