### 2019-02-08 article/tikz/application.tex

\documentclass[border = 1.5cm]{standalone}

%%%% packages
\usepackage{tikz}
\usetikzlibrary{trees,
                arrows,
                shapes.geometric,
                positioning,
                calc,
                backgrounds,
                fit}

%%%% document
\begin{document}

%%%% layers
\pgfdeclarelayer{bg1}    % declare background layer
\pgfdeclarelayer{bg2}
\pgfdeclarelayer{bg3}
\pgfsetlayers{bg3,bg2,bg1,main}  % set the order of the layers (main is the standard layer)

%%%% styles
\tikzset{
	font = {\fontsize{13pt}{12}\sffamily},
	children/.style = {anchor = west},
	main/.style = {rectangle, draw = black, rounded corners, minimum width = 3cm, minimum height = 1.25cm, text centered, anchor = west},
	arrow/.style = {<->, > = stealth, draw = black, color = gray!80, line width = 2.5pt},
	background/.style = {rectangle, rounded corners, minimum width=23cm, fill=gray!10},
}

    %%%% picture
    \begin{tikzpicture}[
	grow via three points={one child at (0.5,-1) and
		two children at (0.5,-1) and (0.5,-1.7)},
	edge from parent path={(\tikzparentnode.south) |- (\tikzchildnode.west)},
	inner sep = 0pt, outer sep = 0pt
	]

    \begin{scope}
	\node (server) [main, at(0cm,0cm), xshift = 4cm, align = center] {Server \\ \textbf{server.R}}
	    child{ node [children, yshift = -0.5cm] { \textbf{setup.R} }}
	    child{ node [children, yshift = -0.5cm] { \textbf{fun\_filter.R} }}
	    child{ node [children, yshift = -0.5cm] { \textbf{fun\_aggregation.R} }};

	\node (ui) [main, right of = server, xshift = 5cm, align = center] {User Interface \\ \textbf{ui.R}};
	\end{scope}

	\begin{pgfonlayer}{bg3}
	\node (back_query) [background, minimum height=5cm, minimum width=14cm, above of=server, xshift=4cm, yshift=-2.25cm];
	\node (back_query_label) [below of = back_query, xshift = 5cm, yshift = -1cm, rotate = 0]
	    {\huge{Application}};
	\end{pgfonlayer}

	\draw [arrow] (server.east) -- (ui.west);

	\end{tikzpicture}
\end{document}