### 2019-06-11 article/parts/04_technical_validation.tex
split up article text

%%This section presents any experiments or analyses that are needed
%%to support the technical quality of the dataset. This section may
%%be supported by up figures and tables, as needed. This is a required
%%section; authors must present information justifying the reliability
%%of their data.

\section{Technical validation}

<PUT COMPARISON WITH PPDB DATA HERE?>