### 2019-09-02 Rmd/summary_db_cols_norman.Rmd
added cbind_fill()

---
date: "`r format(Sys.Date(), '%d.%m.%Y')`"
output:
  html_vignette
  #html_document
editor_options: 
  chunk_output_type: console
params:
  title: "My Title!"
  schema: "information_schema"
  table: "tables"
  output_dir: NULL # https://community.rstudio.com/t/is-it-possible-to-save-the-html-output-in-a-directory-which-is-not-the-one-where-the-rmd-file-resides/3588/6
  cols: NULL
title: "`r params$title`"
---

```{r params, echo=FALSE, include=FALSE}
schema = params$schema
table = params$table
cols = params$cols
dir = file.path(params$output_dir, paste0(schema, '_', table))
# debuging
# schema = 'norman'
# table = 'data1'
# cols = c('nor3', 'nor4')
```

```{r setup, echo=FALSE, include=FALSE}
prj = '/home/scharmueller/Projects/etox-base'
source(file.path(prj, 'R', 'gn_setup.R'))
```

Summary output from data base table __`r paste0(schema, '.', table)`__

```{r echo=FALSE, message=FALSE}
# schema = 'norman'
# table = 'data2_newest'
# cols = c('nor85', 'nor88')

## table
tab = summary_db_all(schema = schema, table = table)
## cols
l = summary_db_cols(schema = schema, table = table, cols = cols)
## names lookup table
q = "SELECT *
     FROM norman.variables"
var = read_tbl(user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
               query = q)
```

```{r preparation, echo=FALSE, message=FALSE}
## table
tab[var, header := i.norman_header, on = c(column = 'id') ]
setcolorder(tab, c('column', 'header', 'type'))
## cols
# TODO names(l) = var[ var$id %in% names(l), norman_header ]
# doesn't work b/c of get(col) in ggplot call
```

```{r function, echo=FALSE}
## plot function
gg_summary = function(l, col, limit, var) {

  dat = l[[col]]
  dat[ , var20 := substr(get(col), 1, 20) ]
  header = var[ match(names(dat)[1], var$id), norman_header ]

  gg = ggplot(dat[ 1:limit ], aes(x = reorder(var20, -n), y = n)) +
    geom_bar(stat = 'identity') +
    geom_text(aes(label = n), position=position_dodge(width=0.9), vjust=-0.25) +
    scale_y_continuous(expand = expand_scale(mult = c(0, 0.05))) +
    labs(title = header, subtitle = col, x = NULL) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

  return(gg)
}
```

## Table
```{r table, echo=FALSE, warning=FALSE}
knitr::kable(tab)
```

## Plots
```{r plots, echo=FALSE, fig.height=7, fig.width=15, dpi=72, warning=FALSE}
# options , dpi=36
mapply(gg_summary,
       col = names(l),
       MoreArgs = list(l = l,
                       limit = 50,
                       var = var),
       SIMPLIFY = FALSE)
```

```{r write, echo=FALSE, message=FALSE}
# write summary to multiple csv
unlink(dir, recursive = TRUE)
mkdirs(dir)
for (i in seq_along(l)) {
  dat = l[[i]]
  nam = names(l)[i]
  fl = file.path(dir, paste0(nam, '.csv'))

  message('Writing: ', fl)
  fwrite(dat, fl)
}
```

```{r write-single-file, message=FALSE}
# write summary to single file
dat = cbind_fill(dflist = l, limit = NULL)
names(dat)[ seq(1, length(names(dat)), 2) + 1 ] = paste0(names(dat)[ seq(1, length(names(dat)), 2) ], '_n')

fwrite(dat, file.path(params$output_dir, paste0(params$table, '_summary.csv')))
```