### 2019-10-14 article/supplement/r-package-list.tex

Rinker TW, Kurkiewicz D (2018).
\emph{pacman: Package Management for R}.
version 0.5.0, \url{http://github.com/trinker/pacman}.
R Core Team (2019).
\emph{R: A Language and Environment for Statistical Computing}.
R Foundation for Statistical Computing, Vienna, Austria.
\url{https://www.R-project.org/}.
Wickham H, Hester J, Chang W (2019).
\emph{devtools: Tools to Make Developing R Packages Easier}.
R package version 2.2.1, \url{https://CRAN.R-project.org/package=devtools}.
Lang DT, team tC (2019).
\emph{RCurl: General Network (HTTP/FTP/...) Client Interface for R}.
R package version 1.95-4.12, \url{https://CRAN.R-project.org/package=RCurl}.
Wickham H (2019).
\emph{stringr: Simple, Consistent Wrappers for Common String Operations}.
R package version 1.4.0, \url{https://CRAN.R-project.org/package=stringr}.
Bengtsson H (2019).
\emph{R.utils: Various Programming Utilities}.
R package version 2.9.0, \url{https://CRAN.R-project.org/package=R.utils}.
Hiebert J (2016).
\emph{udunits2: Udunits-2 Bindings for R}.
R package version 0.13, \url{https://CRAN.R-project.org/package=udunits2}.
Wickham H (2019).
\emph{rvest: Easily Harvest (Scrape) Web Pages}.
R package version 0.3.4, \url{https://CRAN.R-project.org/package=rvest}.
Wickham H (2019).
\emph{httr: Tools for Working with URLs and HTTP}.
R package version 1.4.1, \url{https://CRAN.R-project.org/package=httr}.
Ooms J (2014).
``The jsonlite Package: A Practical and Consistent Mapping Between JSON Data and R Objects.''
\emph{arXiv:1403.2805 [stat.CO]}.
\url{https://arxiv.org/abs/1403.2805}.
Wickham H, Bryan J (2019).
\emph{readxl: Read Excel Files}.
R package version 1.3.1, \url{https://CRAN.R-project.org/package=readxl}.
Walker A (2019).
\emph{openxlsx: Read, Write and Edit XLSX Files}.
R package version 4.1.0.1, \url{https://CRAN.R-project.org/package=openxlsx}.
Henry L, Wickham H (2019).
\emph{purrr: Functional Programming Tools}.
R package version 0.3.2, \url{https://CRAN.R-project.org/package=purrr}.
Dowle M, Srinivasan A (2019).
\emph{data.table: Extension of `data.frame`}.
R package version 1.12.4, \url{https://CRAN.R-project.org/package=data.table}.
Conway J, Eddelbuettel D, Nishiyama T, Prayaga SK, Tiffin N (2017).
\emph{RPostgreSQL: R Interface to the 'PostgreSQL' Database System}.
R package version 0.6-2, \url{https://CRAN.R-project.org/package=RPostgreSQL}.
R Special Interest Group on Databases (R-SIG-DB), Wickham H, Müller K (2018).
\emph{DBI: R Database Interface}.
R package version 1.0.0, \url{https://CRAN.R-project.org/package=DBI}.
Oksanen J, Blanchet FG, Friendly M, Kindt R, Legendre P, McGlinn D, Minchin PR, O'Hara RB, Simpson GL, Solymos P, Stevens MHH, Szoecs E, Wagner H (2019).
\emph{vegan: Community Ecology Package}.
R package version 2.5-5, \url{https://CRAN.R-project.org/package=vegan}.
Wickham H (2011).
``The Split-Apply-Combine Strategy for Data Analysis.''
\emph{Journal of Statistical Software}, \bold{40}(1), 1--29.
\url{http://www.jstatsoft.org/v40/i01/}.
Komsta L (2011).
\emph{outliers: Tests for outliers}.
R package version 0.14, \url{https://CRAN.R-project.org/package=outliers}.
Wickham H (2019).
\emph{feather: R Bindings to the Feather 'API'}.
R package version 0.3.3, \url{https://CRAN.R-project.org/package=feather}.
Wickham H (2016).
\emph{ggplot2: Elegant Graphics for Data Analysis}.
Springer-Verlag New York.
ISBN 978-3-319-24277-4, \url{https://ggplot2.tidyverse.org}.
Slowikowski K (2019).
\emph{ggrepel: Automatically Position Non-Overlapping Text Labels with
'ggplot2'}.
R package version 0.8.1, \url{https://CRAN.R-project.org/package=ggrepel}.
Wilke C (2019).
\emph{cowplot: Streamlined Plot Theme and Plot Annotations for 'ggplot2'}.
R package version 0.9.4, \url{https://CRAN.R-project.org/package=cowplot}.
Neuwirth E (2014).
\emph{RColorBrewer: ColorBrewer Palettes}.
R package version 1.1-2, \url{https://CRAN.R-project.org/package=RColorBrewer}.
Tennekes M (2017).
\emph{treemap: Treemap Visualization}.
R package version 2.4-2, \url{https://CRAN.R-project.org/package=treemap}.
Chamberlain S, Barve V, Mcglinn D, Oldoni D, Desmet P, Geffert L, Ram K (2019).
\emph{rgbif: Interface to the Global Biodiversity Information Facility API}.
R package version 1.3.0, \url{https://CRAN.R-project.org/package=rgbif}.

Chamberlain S, Boettiger C (2017).
``R Python, and Ruby clients for GBIF species occurrence data.''
\emph{PeerJ PrePrints}.
\url{https://doi.org/10.7287/peerj.preprints.3304v1}.
Scott Chamberlain, Eduard Szocs (2013).
``taxize - taxonomic search and retrieval in R.''
\emph{F1000Research}.
\url{http://f1000research.com/articles/2-191/v2}.

Chamberlain S, Szoecs E, Foster Z, Arendsee Z, Boettiger C, Ram K, Bartomeus I, Baumgartner J, O'Donnell J, Oksanen J, Tzovaras BG, Marchand P, Tran V, Salmon M, Li G, Grenié M (2019).
\emph{taxize: Taxonomic information from around the web}.
R package version 0.9.7, \url{https://github.com/ropensci/taxize}.
Arel-Bundock V, Enevoldsen N, Yetman C (2018).
``countrycode: An R package to convert country names and country codes.''
\emph{Journal of Open Source Software}, \bold{3}(28), 848.
\url{https://doi.org/10.21105/joss.00848}.
Microsoft, Weston S (2017).
\emph{foreach: Provides Foreach Looping Construct for R}.
R package version 1.4.4, \url{https://CRAN.R-project.org/package=foreach}.
Corporation M, Weston S (2018).
\emph{doParallel: Foreach Parallel Adaptor for the 'parallel' Package}.
R package version 1.0.14, \url{https://CRAN.R-project.org/package=doParallel}.
Klik M (2019).
\emph{fst: Lightning Fast Serialization of Data Frames for R}.
R package version 0.9.0, \url{https://CRAN.R-project.org/package=fst}.
Xie Y, Cheng J, Tan X (2019).
\emph{DT: A Wrapper of the JavaScript Library 'DataTables'}.
R package version 0.7, \url{https://CRAN.R-project.org/package=DT}.
Tennekes M (2017).
\emph{treemap: Treemap Visualization}.
R package version 2.4-2, \url{https://CRAN.R-project.org/package=treemap}.