### 2019-10-15 article/supplement/r-package-list.tex
article

Rinker TW, Kurkiewicz D (2018).
\emph{pacman: Package Management for R}.
version 0.5.0, \url{http://github.com/trinker/pacman}.
\newline R Core Team (2019).
\emph{R: A Language and Environment for Statistical Computing}.
R Foundation for Statistical Computing, Vienna, Austria.
\url{https://www.R-project.org/}.
\newline Wickham H, Hester J, Chang W (2019).
\emph{devtools: Tools to Make Developing R Packages Easier}.
R package version 2.2.1, \url{https://CRAN.R-project.org/package=devtools}.
\newline Lang DT, team tC (2019).
\emph{RCurl: General Network (HTTP/FTP/...) Client Interface for R}.
R package version 1.95-4.12, \url{https://CRAN.R-project.org/package=RCurl}.
\newline Wickham H (2019).
\emph{stringr: Simple, Consistent Wrappers for Common String Operations}.
R package version 1.4.0, \url{https://CRAN.R-project.org/package=stringr}.
\newline Bengtsson H (2019).
\emph{R.utils: Various Programming Utilities}.
R package version 2.9.0, \url{https://CRAN.R-project.org/package=R.utils}.
\newline Hiebert J (2016).
\emph{udunits2: Udunits-2 Bindings for R}.
R package version 0.13, \url{https://CRAN.R-project.org/package=udunits2}.
\newline Wickham H (2019).
\emph{rvest: Easily Harvest (Scrape) Web Pages}.
R package version 0.3.4, \url{https://CRAN.R-project.org/package=rvest}.
\newline Wickham H (2019).
\emph{httr: Tools for Working with URLs and HTTP}.
R package version 1.4.1, \url{https://CRAN.R-project.org/package=httr}.
\newline Ooms J (2014).
``The jsonlite Package: A Practical and Consistent Mapping Between JSON Data and R Objects.''
\emph{arXiv:1403.2805 [stat.CO]}.
\url{https://arxiv.org/abs/1403.2805}.
\newline Wickham H, Bryan J (2019).
\emph{readxl: Read Excel Files}.
R package version 1.3.1, \url{https://CRAN.R-project.org/package=readxl}.
\newline Walker A (2019).
\emph{openxlsx: Read, Write and Edit XLSX Files}.
R package version 4.1.0.1, \url{https://CRAN.R-project.org/package=openxlsx}.
\newline Henry L, Wickham H (2019).
\emph{purrr: Functional Programming Tools}.
R package version 0.3.2, \url{https://CRAN.R-project.org/package=purrr}.
\newline Dowle M, Srinivasan A (2019).
\emph{data.table: Extension of `data.frame`}.
R package version 1.12.4, \url{https://CRAN.R-project.org/package=data.table}.
\newline Conway J, Eddelbuettel D, Nishiyama T, Prayaga SK, Tiffin N (2017).
\emph{RPostgreSQL: R Interface to the 'PostgreSQL' Database System}.
R package version 0.6-2, \url{https://CRAN.R-project.org/package=RPostgreSQL}.
\newline R Special Interest Group on Databases (R-SIG-DB), Wickham H, Müller K (2018).
\emph{DBI: R Database Interface}.
R package version 1.0.0, \url{https://CRAN.R-project.org/package=DBI}.
\newline Oksanen J, Blanchet FG, Friendly M, Kindt R, Legendre P, McGlinn D, Minchin PR, O'Hara RB, Simpson GL, Solymos P, Stevens MHH, Szoecs E, Wagner H (2019).
\emph{vegan: Community Ecology Package}.
R package version 2.5-5, \url{https://CRAN.R-project.org/package=vegan}.
\newline Wickham H (2011).
``The Split-Apply-Combine Strategy for Data Analysis.''
\emph{Journal of Statistical Software}, \bold{40}(1), 1--29.
\url{http://www.jstatsoft.org/v40/i01/}.
\newline Komsta L (2011).
\emph{outliers: Tests for outliers}.
R package version 0.14, \url{https://CRAN.R-project.org/package=outliers}.
\newline Wickham H (2019).
\emph{feather: R Bindings to the Feather 'API'}.
R package version 0.3.3, \url{https://CRAN.R-project.org/package=feather}.
\newline Wickham H (2016).
\emph{ggplot2: Elegant Graphics for Data Analysis}.
Springer-Verlag New York.
ISBN 978-3-319-24277-4, \url{https://ggplot2.tidyverse.org}.
\newline Slowikowski K (2019).
\emph{ggrepel: Automatically Position Non-Overlapping Text Labels with
'ggplot2'}.
R package version 0.8.1, \url{https://CRAN.R-project.org/package=ggrepel}.
\newline Wilke C (2019).
\emph{cowplot: Streamlined Plot Theme and Plot Annotations for 'ggplot2'}.
R package version 0.9.4, \url{https://CRAN.R-project.org/package=cowplot}.
\newline Neuwirth E (2014).
\emph{RColorBrewer: ColorBrewer Palettes}.
R package version 1.1-2, \url{https://CRAN.R-project.org/package=RColorBrewer}.
\newline Tennekes M (2017).
\emph{treemap: Treemap Visualization}.
R package version 2.4-2, \url{https://CRAN.R-project.org/package=treemap}.
\newline Chamberlain S, Barve V, Mcglinn D, Oldoni D, Desmet P, Geffert L, Ram K (2019).
\emph{rgbif: Interface to the Global Biodiversity Information Facility API}.
R package version 1.3.0, \url{https://CRAN.R-project.org/package=rgbif}.

Chamberlain S, Boettiger C (2017).
``R Python, and Ruby clients for GBIF species occurrence data.''
\emph{PeerJ PrePrints}.
\url{https://doi.org/10.7287/peerj.preprints.3304v1}.
\newline Scott Chamberlain, Eduard Szocs (2013).
``taxize - taxonomic search and retrieval in R.''
\emph{F1000Research}.
\url{http://f1000research.com/articles/2-191/v2}.

Chamberlain S, Szoecs E, Foster Z, Arendsee Z, Boettiger C, Ram K, Bartomeus I, Baumgartner J, O'Donnell J, Oksanen J, Tzovaras BG, Marchand P, Tran V, Salmon M, Li G, Grenié M (2019).
\emph{taxize: Taxonomic information from around the web}.
R package version 0.9.7, \url{https://github.com/ropensci/taxize}.
\newline Arel-Bundock V, Enevoldsen N, Yetman C (2018).
``countrycode: An R package to convert country names and country codes.''
\emph{Journal of Open Source Software}, \bold{3}(28), 848.
\url{https://doi.org/10.21105/joss.00848}.
\newline Microsoft, Weston S (2017).
\emph{foreach: Provides Foreach Looping Construct for R}.
R package version 1.4.4, \url{https://CRAN.R-project.org/package=foreach}.
\newline Corporation M, Weston S (2018).
\emph{doParallel: Foreach Parallel Adaptor for the 'parallel' Package}.
R package version 1.0.14, \url{https://CRAN.R-project.org/package=doParallel}.
\newline Klik M (2019).
\emph{fst: Lightning Fast Serialization of Data Frames for R}.
R package version 0.9.0, \url{https://CRAN.R-project.org/package=fst}.
\newline Xie Y, Cheng J, Tan X (2019).
\emph{DT: A Wrapper of the JavaScript Library 'DataTables'}.
R package version 0.7, \url{https://CRAN.R-project.org/package=DT}.
\newline Tennekes M (2017).
\emph{treemap: Treemap Visualization}.
R package version 2.4-2, \url{https://CRAN.R-project.org/package=treemap}.
\newline