### 2019-10-15 article/00_article.tex

\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage{natbib} % citations
\usepackage{csvsimple}
\usepackage{pgfplotstable}
\usepackage{rotating} % to rotate tables
%% tikz packages
\usepackage{tikz}
\usepackage{forest}
\usetikzlibrary{trees, arrows, shapes.geometric, positioning}
\usepackage{standalone}
%% text packages
\usepackage{siunitx} % fake ° symbol \ang{}
\usepackage[super]{nth} % for 1st, 2nd etc. \nth{}
\usepackage{hyperref}

\newcommand{\standartox}{STANDARTOX}
\newcommand{\epa}{EPA ECOTOX data base}
\newcommand{\rpackage}{http://139.14.20.252:3838/etox-base-shiny/}
\newcommand{\git}{https://github.com/andreasLD/etox-base}
\newcommand{\gitapp}{https://github.com/andreasLD/etox-base-shiny}
\newcommand{\ecfifty}{EC\textsubscript{50}}

\listfiles

\begin{document}

\title{Standartox: A tool to derive exposure endpoints from chemical test results}

\author{Andreas Scharm{\"u}ller\textsuperscript{1{*}},
        Verena Schreiner\textsuperscript{1},
        Ralf B. Sch{\"a}fer\textsuperscript{1}}

\maketitle
\thispagestyle{fancy}

1. Institute of Environmental Sciences, University of Koblenz-Landau Fortstraße 7, 76829 Landau, Germany {*}corresponding author(s):
Andreas Scharm{\"u}ller (andschar@protonmail.com)

\begin{abstract}
%%5This is a manuscript template for Data Descriptor submissions to \emph{Scientific
%%Data} (http://www.nature.com/scientificdata). The abstract must be
%%no longer than 170 words, and should succinctly describe the study,
%%the assay(s) performed, the resulting data, and the reuse potential,
%%but should not make any claims regarding new scientific findings.
%%No references are allowed in this section.

<WRITE ABSTRACT>

\end{abstract}
\pagebreak

%%%%%%%%%%%%%%%%%%%%%% 01 Background
\input{article/parts/01_background.tex}
\pagebreak

%%%%%%%%%%%%%%%%%%%%%% 02 Methods
\input{article/parts/02_methods.tex}
\pagebreak

%%%%%%%%%%%%%%%%%%%%%% 03 Data records
\input{article/parts/03_data_records.tex}
\pagebreak

%%%%%%%%%%%%%%%%%%%%%% 04 Technical validation
\input{article/parts/04_technical_validation.tex}
\pagebreak

%%%%%%%%%%%%%%%%%%%%%% 05 Usage notes
\input{article/parts/05_usage_notes.tex}

\section*{Acknowledgements}
%%Text acknowledging non-author contributors. Acknowledgements should
%%be brief, and should not include thanks to anonymous referees and
%%editors, or effusive comments. Grant or contribution numbers may be
%%acknowledged. Author contributions Please describe briefly the contributions
%%of each author to this work on a separate line. 
%%
%%AK did this and that. 
%%
%%BG did this and that and the other.

\section*{Competing financial interests}
%%A competing financial interests statement is required for all accepted
%%papers published in \emph{Scientific Data}. If none exist simply write,
%%``The author(s) declare no competing financial interests''.

\section*{Figures Legends}
%%Figure should be referred to using a consistent numbering scheme through
%%the entire Data Descriptor. For initial submissions, authors may choose
%%to supply this document as a single PDF with embedded figures, but
%%separate figure image files must be provided for revisions and accepted
%%manuscripts. In most cases, a Data Descriptor should not contain more
%%than three figures, but more may be allowed when needed. We discourage
%%the inclusion of figures in the Supplementary Information \textendash{}
%%all key figures should be included here in the main Figure section. 
%%
%%Figure legends begin with a brief title sentence for the whole figure
%%and continue with a short description of what is shown in each panel,
%%as well as explaining any symbols used. Legend must total no more
%%than 350 words, and may contain literature references.

\section*{Tables}
%%Tables supporting the Data Descriptor. These can provide summary information
%%(sample numbers, demographics, etc.), but they should generally not
%%be used to present primary data (i.e. measurements). Tables containing
%%primary data should be submitted to an appropriate data repository. 
%%
%%Tables may be provided within the \LaTeX{} document or as separate
%%files (tab-delimited text or Excel files). Legends, where needed,
%%should be included here. Generally, a Data Descriptor should have
%%fewer than ten Tables, but more may be allowed when needed. Tables
%%may be of any size, but only Tables which fit onto a single printed
%%page will be included in the PDF version of the article (up to a maximum
%%of three).

%\begin{sidewaystable}
%    \csvautotabular[respect all]{data/meta.csv};
%    \caption{Meta variables table};
%    \label{table:meta-variables};
%\end{sidewaystable}

\pagebreak

\bibliographystyle{apalike}
\bibliography{refs/references-standartox.bib} % only reference from Zotero to avoid key conflicts
% NOTE multiple refs would have to be written w/o space and separated by <,>

\section*{Data Citations}

\pagebreak
\section*{TODO}

\begin{itemize}

\item compare Etox-BAse output with PPDB

\item plot example SSD and TU

\item acute vs chronic

\item OECD guidlines are not machine readable

\item discuss flaws in column types (e.g some concentration entries contain '+' or '~' and are therefore not convertable to numeric) - This hampers large scale analysis greatly!

\pagebreak
\section*{MISC}

\begin{figure}
    \includestandalone[scale=0.5]{tikz/tree_source}
    \caption{Alternative processing pipeline}
    \label{fig:pipeline-tree}
\end{figure}

\section*{not included citations}

\item Put in citation \citep{hartung_chemical_2009} ? which claims that REACH won't meet their assumptions and say 54 million vertebrate animals are needed for tests that would cost €9.5 billion over the next ten years. I.e 20 times more animals, 6 times the costs in comparison to the official estimates

\item "comparison of rate of change in the production and variety of pesticides, pharmaceuticals and other chemicals over the past four decades to CO2 nutrient pollution and other global change factors" \citep{bernhardt_synthetic_2017}

\item worldwide decline in entomofauna: \citep{sanchez-bayo_worldwide_2019}

\end{itemize}

\pagebreak
\section*{Supplement}
\subsection*{R-package list} %% TODO put in appropriate list
\input{article/supplement/r_package_list.tex}
\pagebreak

\subsection*{Data base tables}
\input{article/supplement/data_base_tables.tex}

\end{document}