### 2019-10-18 article/tables/DEPRECATED_old_tables.tex
Updates from Overleaf

\begin{table}[h!]
\begin{tabular}{|l|l|}
\hline
Script                      & Description                                      \\ \hline
run\_build.R                 & runs all sub scripts                             \\ \hline
setup.R                     & sets up the environment                          \\ \hline
bd\_epa\_download.R         & download EPA Ecotox data base                    \\ \hline
bd\_epa\_postgres.R         & build EPA Ecotox data base locally               \\ \hline
da\_epa1.R                  & raw EPA Ecotox data export                       \\ \hline
da\_epa2.R                  & process EPA Ecotox data                          \\ \hline
da\_epa3.R                  & reduce EPA Ecotox data                           \\ \hline
qu\_aw.R                    & query Compendium of Pesticide Common Names       \\ \hline
qu\_chemspider\_scrape.R    & query Chemspider data base                       \\ \hline
qu\_eurostat\_chem\_class.R & query Eurostat data set                          \\ \hline
qu\_pc.R                    & query PubChem data base                          \\ \hline
qu\_pp.R                    & query Physprop data base                         \\ \hline
qu\_worms2.R                & query WORMS data base                            \\ \hline
qu\_gbif.R                  & query GBIF data base                             \\ \hline
re\_merge.R                 & merge all data                                   \\ \hline
re\_combine.R               & combine information                              \\ \hline
re\_analyses.R              & ???                                              \\ \hline
re\_checks\_internal.R      & ???                                              \\ \hline
re\_final.R                 & create and write final table                     \\ \hline
wr\_dbase.R                 & write final table to data base                   \\ \hline
wr\_app.R                   & write final table and variables to app directory \\ \hline
wr\_meta.R                  & ???                                              \\ \hline
wr\_stats.R                 & ???                                              \\ \hline
\end{tabular}
\caption{Scripts that are used to process the data}
\label{tab:scripts-pipeline}
\end{table}

\begin{table}[h!]
\begin{tabular}{|l|l|}
\hline
Script                     & Description                                               \\ \hline
setup.R                    & sets up the environment                                   \\ \hline
ui.R                       & creates the graphical user interface of the application   \\ \hline
server.R                   & handles functions and data input and output               \\ \hline
fun\_filter.R              & function that filters the data                            \\ \hline
fun\_aggregation.R         & function that performs the aggregation                    \\ \hline
fun\_filagg\_plot\_ly.R    & creates interactive plot                                  \\ \hline
\end{tabular}
\caption{Scripts that used for the web application}
\label{tab:scripts-app}
\end{table}