### 2019-10-18 article/tables/data_base_tables.tex

\begin{table}[ht]
  \caption{Table of queried data bases}
  \label{tab:data-base}
  \centering
\begin{tabular}{ll}
  \hline
  Data Base & URL \\ 
  \hline
  Chemical Entities of Biological Interest (ChEBI) & https://www.ebi.ac.uk/chebi \\ 
  Chemical Identifier Resolver & https://cactus.nci.nih.gov/chemical/structure \\
  ChemSpider & http://www.chemspider.com \\
  Eurostat & https://ec.europa.eu/eurostat/home \\
  PHYSPROP Database & https://www.srcinc.com/what-we-do/environmental/scientific-databases.html \\
  PubChem & https://pubchem.ncbi.nlm.nih.gov \\ 
  Wikidata & https://www.wikidata.org/wiki/Wikidata:Main_Page \\
  Global Biodiversity Information Facility (GBIF) & https://www.gbif.org \\
  World Register of Marine Species (WoRMS) & http://marinespecies.org \\
  \hline
\end{tabular}
\end{table}