### 2019-10-18 article/parts/04_technical_validation.tex

%%This section presents any experiments or analyses that are needed
%%to support the technical quality of the dataset. This section may
%%be supported by up figures and tables, as needed. This is a required
%%section; authors must present information justifying the reliability
%%of their data.

\section{Technical validation}

\begin{figure}
    \includegraphics[width=1\linewidth]{article/figures/.png}
    \caption{
        Comparison between Standartox and PPDB values
    }
    \label{fig:standartox_ppdb_diff}
\end{figure}

\subsection{PPDB Comparsion}
The PPDB also offers effect values.

\subsection{Unit testing}

\subsection{Overall distribution of tests}
Heatmap: don't Include for now. Too large

In order to validate Standartox we compared resulting values with values from other data bases, such as the PPDB. Of 542 chemicals, only 16\% diverged from the PPDB values by a factor greater than 10, supporting plausibility of the \standartox results \ref{fig:standartox_ppdb_diff}.

<PUT COMPARISON WITH PPDB DATA HERE?>