### 2019-11-05 article/parts/06_supplement.tex

\section*{Supplement}
\subsection*{R-package list} %% TODO put in appropriate list
\label{list:r-packages}
%% TODO how to refernce this?
\input{article/tables/r_package_list.tex}
\pagebreak

\subsection*{Data base tables}
\input{article/tables/data_base_tables.tex}

\subsection*{Unit conversion}

\begin{table}
    \csvautotabular[respect all]{lookup/lookup_concentration_unit.csv}
    \caption{Concentration unit conversion}
    \label{tab:conv-concentration}
\end{table}

\begin{table}
    \csvautotabular[respect all]{lookup/lookup_duration.csv}
    \caption{Duration unit conversion}
    \label{tab:conv-duration}
\end{table}

\subsection*{Application parameters}
\input{article/tables/app_parameters.tex}

\input{article/tables/chck-units-concentration.tex}

\input{article/tables/chck-units-duration.tex}