### 2019-11-06 article/tables/look-concentration.units.tex
added latex table

\begin{tabular}{l|l|l|l|r|l|l|l|l|l}
\hline
unit & prefix & word & si & multiplier & conv & unit\_conv & type & remove & notes\\
\hline
fg & femto & gram & yes & 0.000000e+00 & yes & g & mass & no & \\
\hline
pg & pico & gram & yes & 0.000000e+00 & yes & g & mass & no & \\
\hline
ng & nano & gram & yes & 0.000000e+00 & yes & g & mass & no & \\
\hline
ug & micro & gram & yes & 1.000000e-06 & yes & g & mass & no & \\
\hline
mg & milli & gram & yes & 1.000000e-03 & yes & g & mass & no & \\
\hline
g &  & gram & yes & 1.000000e+00 & yes & g & mass & no & \\
\hline
kg & kilo & gram & yes & 1.000000e+03 & yes & g & mass & no & \\
\hline
t & mega & gram & no & 1.000000e+06 & yes & g & mass & no & \\
\hline
tons & mega & gram & no & 1.000000e+06 & yes & g & mass & no & \\
\hline
pl & pico & liter & yes & 0.000000e+00 & yes & l & volume & no & \\
\hline
nl & nano & liter & yes & 0.000000e+00 & yes & l & volume & no & \\
\hline
ul & micro & liter & yes & 1.000000e-06 & yes & l & volume & no & \\
\hline
ml & milli & liter & yes & 1.000000e-03 & yes & l & volume & no & \\
\hline
l &  & liter & yes & 1.000000e+00 & yes & l & volume & no & \\
\hline
hl & hecto & liter & yes & 1.000000e+02 & yes & l & volume & no & \\
\hline
bu &  & bushel & no & 3.523907e+01 & yes & l & volume & no & \\
\hline
ppq & quadrillion & partsper & no & 1.000000e-06 & yes & ppb & fraction & no & \\
\hline
ppt & trillion & partsper & no & 1.000000e-03 & yes & ppb & fraction & no & \\
\hline
ppb & billion & partsper & no & 1.000000e+00 & yes & ppb & fraction & no & \\
\hline
ppm & million & partsper & no & 1.000000e+03 & yes & ppb & fraction & no & \\
\hline
no &  & number & no & NA & no &  & amount & no & \\
\hline
amol & atto & mol & yes & 0.000000e+00 & yes & mol & mol & no & \\
\hline
fmol & femto & mol & yes & 0.000000e+00 & yes & mol & mol & no & \\
\hline
pmol & pico & mol & yes & 0.000000e+00 & yes & mol & mol & no & \\
\hline
nmol & nano & mol & yes & 0.000000e+00 & yes & mol & mol & no & \\
\hline
umol & micro & mol & yes & 1.000000e-06 & yes & mol & mol & no & \\
\hline
mmol & milli & mol & yes & 1.000000e-03 & yes & mol & mol & no & \\
\hline
mol & one & mol & yes & 1.000000e+00 & yes & mol & mol & no & \\
\hline
kmol & kilo & mol & yes & 1.000000e+03 & yes & mol & mol & no & \\
\hline
nM & nano & mol & yes & 0.000000e+00 & yes & mol & mol & no & \\
\hline
uM & micro & mol & yes & 1.000000e-06 & yes & mol & mol & no & \\
\hline
mM & milli & mol & yes & 1.000000e-03 & yes & mol & mol & no & \\
\hline
M & one & mol & yes & 1.000000e+00 & yes & mol & mol & no & \\
\hline
yd &  & yard & no & 9.144000e-01 & yes & m & length & no & \\
\hline
ft &  & foot & no & 3.048000e-01 & yes & m & length & no & \\
\hline
cm & centi & meter & yes & 1.000000e-02 & yes & m & length & no & \\
\hline
mm & milli & meter & yes & NA & yes &  & length &  & \\
\hline
m &  & meter & yes & NA &  &  & length &  & \\
\hline
neq & nano & equivalent & no & NA & no &  & noscience & yes & \\
\hline
meq & milli & equivalent & no & NA & no &  & noscience & yes & \\
\hline
ueq & micro & equivalent & no & NA & no &  & noscience & yes & \\
\hline
mm2 & milli &  & yes & 1.000000e-03 & yes & m2 & area & no & \\
\hline
cm2 & centi &  & yes & 1.000000e-02 & yes & m2 & area & no & \\
\hline
m2 & one &  & yes & 1.000000e+00 & yes & m2 & area & no & \\
\hline
ha &  & hectare & yes & 1.000000e+04 & yes & m2 & area & no & \\
\hline
acre &  &  & no & 4.050000e+03 & yes & m2 & area & no & \\
\hline
ac &  &  & no & 4.050000e+03 & yes & m2 & area & no & \\
\hline
mi2 &  & squaremile & no & NA & yes & m2 & area & no & \\
\hline
km2 & kilo & squaremeter & yes & 1.000000e+06 & yes & m2 & area & no & \\
\hline
ft2 &  & squarefeet & no & 9.290000e-02 & yes & m2 & area & no & \\
\hline
feddan &  & feddan & no & 4.200000e+03 & yes & m2 & area & no & \\
\hline
dn(Cyprus) &  & donum & no & 3.300000e+02 & yes & m2 & area & no & \\
\hline
\% &  & percent & no & 1.000000e+07 & yes & ppb & percentage & no & \\
\hline
‰ &  & permille & no & 1.000000e+06 & yes & ppb & percentage & no & \\
\hline
d &  & day & no & 2.400000e+01 & yes & h & time & no & \\
\hline
h &  & hour & no & 1.000000e+00 & yes & h & time &  & \\
\hline
mi &  & minute & no & NA &  &  & time &  & \\
\hline
wk &  & week & no & NA &  &  & time &  & \\
\hline
yr &  & year & no & NA & no &  & time &  & \\
\hline
ft3 &  & squarefeet & no & NA &  &  & volume & no & \\
\hline
mm3 & milli & cubic & no & 0.000000e+00 & yes & m3 & volume & no & \\
\hline
cm3 & centi & cubic & no & 1.000000e-06 & yes & m3 & volume & no & \\
\hline
dm3 & deci & cubic & no & 1.000000e-03 & yes & m3 & volume & no & \\
\hline
m3 &  & cubic & no & 1.000000e+00 & yes & m3 & volume & no & \\
\hline
fl oz &  &  & no & 2.960000e-02 & yes & l & volume & no & 28.4ml in the imperial system – 29.6ml in the US\\
\hline
pt &  & pint & no & NA & no &  & volume & yes & three different pints\\
\hline
oz &  & ounce & no & 2.834000e+01 & yes & g & volume & no & \\
\hline
gal &  &  & no & 4.546090e+00 & yes & l & volume & no & \\
\hline
qt &  & quarts & no & NA &  &  & volume &  & \\
\hline
lb &  & pound & no & 4.535920e+02 & yes & g & mass & no & \\
\hline
psu &  & practical salinity unit & no & NA & no &  & salinity & no & \\
\hline
-- &  & no data & no & NA & no &  & nodata & yes & \\
\hline
nr &  & no data & no & NA & no &  & nodata & yes & \\
\hline
ai &  & active ingredient & no & NA & no &  & noscience & yes & \\
\hline
eu &  & experimental unit & no & 1.000000e+00 & no & eu & experimentalunit & yes & \\
\hline
kbq & kilo & becquerel & no & NA &  &  & radioactivity &  & \\
\hline
bq &  & becquerel & no & NA & no &  & radioactivity & no & \\
\hline
mbq & milli & becquerel &  & NA &  &  & radioactivity &  & \\
\hline
ubq & micro & becquerel & no & NA & no &  & radioactivity & yes & \\
\hline
ci &  & curie & no & 0.000000e+00 & yes & Bq & radioactivity & no & \\
\hline
mci & milli & curie &  & NA &  &  & radioactivity &  & \\
\hline
uci & micro & curie & no & NA &  &  & radioactivity &  & \\
\hline
nci & nano & curie &  & NA &  &  & radioactivity &  & \\
\hline
icu &  & international chick units & no & NA & no &  & noscience & yes & \\
\hline
iu &  & international units & no & NA & no &  & noscience & yes & \\
\hline
fibers &  & fibers & no & NA & no &  & noscience & yes & \\
\hline
kj & kilo & joule & no & NA & no &  & energy & yes & \\
\hline
n &  & normal & no & NA & no &  & noscience & yes & \\
\hline
ra &  & ratio & no & NA & no &  & ratio & yes & \\
\hline
unit &  & unit & no & NA & no &  & noscience & yes & \\
\hline
org &  & organism & no & 1.000000e+00 & no & org & organism & yes & \\
\hline
v &  & volume & no & NA & no &  & volume & yes & \\
\hline
\% v &  & percentage volume &  & NA &  &  & percentage &  & \\
\hline
cwt &  & hundred weight & no & NA & no &  & noscience & yes & \\
\hline
w &  & weight & no & NA & no &  & noscience & yes & \\
\hline
\% w &  & percentage weight &  & NA &  &  & percentage &  & \\
\hline
in dia &  & inch diameter & no & NA & no &  & noscience & yes & \\
\hline
egg &  & egg & no & NA & no &  & noscience & yes & \\
\hline
pellets &  & pellets & no & NA & no &  & noscience & yes & \\
\hline
bee &  & bee & no & NA & no &  & noscience & yes & \\
\hline
fish &  & fish & no & NA & no &  & noscience & yes & \\
\hline
u &  & units & no & NA & no &  & noscience & yes & \\
\hline
dpm &  & disintegrations per minute & no & NA & no &  & radioactivity & yes & \\
\hline
cwt &  & hundredweight & no & NA & no &  & noscience & yes & \\
\hline
sd &  & seed & no & NA & no &  & noscience & yes & \\
\hline
cntr &  & container &  & NA &  &  & noscience &  & \\
\hline
plot &  & plot &  & NA &  &  & noscience &  & \\
\hline
cpm &  & counts per minute &  & NA &  &  & counts &  & \\
\hline
\end{tabular}