### 2019-11-18 article/tables/app-parameters.tex

\begin{table}[ht]
  \caption{Input parameters for the Standartox Web application and the R-package standartox.}
  \label{tab:app-parameters}
  \centering
\begin{tabular}{ll}
  \hline
  parameter & example \\ 
  \hline
  cas & 7758987, 2921882, 1912249 \\
  concentration\_type & active ingredient, formulation, total \\
  chemical\_class & fungicide, herbicide, insecticide \\
  taxa & Oncorhynchus mykiss, Rattus norvegicus, Daphnia magna \\
  habitat & marine, brackish, freshwater \\
  region & africa, america_north, america_south \\
  duration & 24, 96 \\
  effect & Mortality, Population, Growth \\
  endpoint & NOEX, XX50, LOEX \\
  vers & 20190912 \\
  \hline
\end{tabular}
\end{table}