### 2019-11-18 article/00_article.tex

\documentclass[english]{article}

%% ---------------------------- packages ------------------------------------ %%
\usepackage[T1]{fontenc}
\usepackage{natbib} % citations
\usepackage{csvsimple, longtable, booktabs}
\usepackage{pgfplotstable}
\usepackage{rotating} % to rotate tables
%% tikz packages
\usepackage{tikz}
\usepackage{forest}
\usetikzlibrary{trees, arrows, shapes.geometric, positioning}
\usepackage{standalone}
%% text packages
\usepackage{siunitx} % fake ° symbol \ang{}
\usepackage[super]{nth} % for 1st, 2nd etc. \nth{}
\usepackage{hyperref}
\usepackage{listings} % code highlighting
\usepackage{color}

%% ---------------------------- commands ------------------------------------ %%
\newcommand{\standartox}{STANDARTOX}
\newcommand{\epa}{EPA ECOTOX data base}
\newcommand{\rpackage}{http://139.14.20.252:3838/etox-base-shiny/}
\newcommand{\git}{https://github.com/andreasLD/etox-base}
\newcommand{\gitapp}{https://github.com/andreasLD/etox-base-shiny}
\newcommand{\ecfifty}{EC\textsubscript{50}}

%% ---------------------------- code ---------------------------------------- %%
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\lstset{frame=tb,
  language=R,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=true,
  tabsize=3
}

\listfiles

\begin{document}

\title{Standartox: A tool to derive exposure endpoints from ecotoxicological test results}

\author{Andreas Scharm{\"u}ller\textsuperscript{1{*}},
        Verena C. Schreiner\textsuperscript{1},
        Ralf B. Sch{\"a}fer\textsuperscript{1}}

\maketitle
\thispagestyle{fancy}

1. Institute of Environmental Sciences, University of Koblenz-Landau Fortstraße 7, 76829 Landau, Germany {*}corresponding author(s):
Andreas Scharm{\"u}ller (andschar@protonmail.com)

\begin{abstract}
%%5This is a manuscript template for Data Descriptor submissions to \emph{Scientific
%%Data} (http://www.nature.com/scientificdata). The abstract must be
%%no longer than 170 words, and should succinctly describe the study,
%%the assay(s) performed, the resulting data, and the reuse potential,
%%but should not make any claims regarding new scientific findings.
%%No references are allowed in this section.

<WRITE ABSTRACT>

\end{abstract}
\pagebreak

%%%%%%%%%%%%%%%%%%%%%% 01 Introduction
\input{article/parts/01_introduction.tex}

%%%%%%%%%%%%%%%%%%%%%% 02 Results and Discussion
\input{article/parts/02_results_discussion.tex}

%%%%%%%%%%%%%%%%%%%%%% 03 Methods
\input{article/parts/03_methods.tex}

%%%%%%%%%%%%%%%%%%%%%% 04 Conclusion
\input{article/parts/04_conclusion.tex}

%%%%%%%%%%%%%%%%%%%%%% 05 Code Availability
\input{article/parts/05_code_availability.tex}

%%%%%%%%%%%%%%%%%%%%%% References
% only reference from Zotero to avoid key conflicts
% NOTE multiple refs would have to be written w/o space and separated by <,>
\bibliographystyle{apalike}
\bibliography{refs/references-standartox.bib}

\section*{Acknowledgements}
%%Text acknowledging non-author contributors. Acknowledgements should
%%be brief, and should not include thanks to anonymous referees and
%%editors, or effusive comments. Grant or contribution numbers may be
%%acknowledged. Author contributions Please describe briefly the contributions
%%of each author to this work on a separate line. 
%%
%%AK did this and that. 
%%
%%BG did this and that and the other.

\section*{Competing financial interests}
%%A competing financial interests statement is required for all accepted
%%papers published in \emph{Scientific Data}. If none exist simply write,
%%``The author(s) declare no competing financial interests''.

\section*{Author Information}

\section*{Ethics Declaration}

\section*{Additional Information}

\section*{Rights and permissions}

%%%%%%%%%%%%%%%%%%%%%% 10 Supplement
\input{article/parts/10_supplement.tex}

\end{document}