### 2019-11-18 article/parts/10_supplement.tex

\section*{Supplement}

\subsection*{R-package list}
\label{list:r-packages}
\input{article/tables/r_package_list.tex}

\subsection*{Unit conversions}
 TODO
\begin{table}
    \csvautotabular[respect all]{lookup/lookup_concentration_unit.csv}
    \caption{Concentration unit conversion}
    \label{sup:conv-concentration}
\end{table}

\begin{table}
    \csvautotabular[respect all]{lookup/lookup_duration.csv}
    \caption{Duration unit conversion}
    \label{sup:conv-duration}
\end{table}

%\input{article/tables/chck-units-concentration.tex}

%\input{article/tables/chck-units-duration.tex}

\subsection*{Scripts}
\begin{figure}
    % TODO make better graphic
    \includegraphics[width=0.5\linewidth]{article/figures/scripts_dendrogram.png}
    \caption{Processing pipeline}
    \label{fig:pipeline-tree}
\end{figure}