### 2019-11-22 article/tables/app-parameters.tex

\begin{table}[ht]
  \caption{Input parameters for the Standartox Web application and the R-package standartox.}
  \label{tab:app-parameters}
  \centering
\begin{tabular}{cc}
  \hline
  parameter & example \\ 
  \hline
  cas & 7758987, 2921882, 1912249   \\[0.5cm]
  concentration\_type & active ingredient, formulation, total   \\[0.5cm]
  chemical\_class & fungicide, herbicide, insecticide   \\[0.5cm]
  taxa & Oncorhynchus mykiss, Rattus norvegicus, Daphnia magna   \\[0.5cm]
  habitat & marine, brackish, freshwater   \\[0.5cm]
  region & africa, america\_north, america\_south   \\[0.5cm]
  duration & 24, 96   \\[0.5cm]
  effect & Mortality, Population, Growth   \\[0.5cm]
  endpoint & NOEX, XX50, LOEX   \\[0.5cm]
  vers & 20190912   \\
  \hline
\end{tabular}
\end{table}