### 2019-11-28 article/code/standartox_example.tex

\begin{lstlisting}[
    caption = {Sample code to access the Standartox data base through the API with the help of the R-package. \textit{stx\_catalog()} returns a catalog of possible filter and aggregation parameters. \textit{stx\_query()} returns the Standartox object, a list of the filtered and the aggregated data as well as a meta data entry.},
    label = {listing:standartox-example}]
# install
install.packages('remotes')
remotes::install_github('andschar/standartox') # package not yet on CRAN
# retrieve catalog    
require(standartox)
catal = stx_catalog()
# retrieve data
l = stx_query(cas = '1071-83-6',
                endpoint = 'XX50',
                taxa = 'Oncorhynchus',
                duration = c(24, 120))
\end{lstlisting}