### 2019-12-16 article/parts/04_user_notes.tex
Updates from Overleaf

\section{User notes}

Using the APP, the user can download the resulting data sets as a comma-separated values (csv) file whereas by using the API, the data can directly be loaded in R.  To provide an easy overview of the data set, percentage values for each parameter entry are shown in the APP in the selection menu as well as by using the R package through querying the \textit{/catalog} API-endpoint.

\pagebreak

\begin{figure}
    \includestandalone[scale=0.4]{article/tikz/standartox_organigram}
    \caption{Organigram of Standartox.}
    \label{fig:stx-organigram}
\end{figure}