### 2019-12-16 article/parts/05_conclusion.tex

\section{Conclusion}

Standartox is a tool that does this in a generalizing way and puts its focus on the aggregation of toxicity data, by this adding a piece to the puzzle of modern ecotoxicological analyses.