### 2019-12-16 article/tables/data_base_tables.tex

\begin{table}[ht]
  \caption{Table of additionally queried publicly available data bases and their URLs.}
  \label{tab:data-base-additional}
  \centering
\begin{tabular}{cc}
  \hline
  Data Base & URL \\ 
  \hline
  \makecell{Chemical Entities of \\ Biological Interest (ChEBI)} & https://www.ebi.ac.uk/chebi  \\
  Chemical Identifier Resolver & https://cactus.nci.nih.gov/chemical/structure  \\[0.5cm]
  ChemSpider & http://www.chemspider.com    \\[0.5cm]
  Eurostat & https://ec.europa.eu/eurostat/home \\[0.5cm]
  PHYSPROP & \makecell{https://www.srcinc.com/what-we-do/environmental/\\scientific-databases.html}  \\[0.5cm]
  PubChem & \url{https://pubchem.ncbi.nlm.nih.gov} \\[0.5cm]
  Wikidata & \url{https://www.wikidata.org/wiki/Wikidata:Main_Page} \\[0.5cm]
  \makecell{Global Biodiversity \\ Information Facility (GBIF)} & https://www.gbif.org \\[0.5cm]
  \makecell{World Register of \\ Marine Species (WoRMS)} & http://marinespecies.org \\
  \hline
\end{tabular}
\end{table}