### 2019-12-16 article/tikz/pipeline-organigram-detailed.tex

\documentclass[border = 1.5cm]{standalone}

\usepackage{tikz}
\usetikzlibrary{trees,
                arrows,
                shapes.geometric,
                positioning,
                calc,
                backgrounds,
                fit}
%% document
\begin{document}

\tikzset{
    font = {\fontsize{13pt}{12}\sffamily},
    epa/.style = {rectangle, text centered, draw=black, fill=blue!10, minimum width=3cm, minimum height=3cm},
    script-r/.style = {rectangle, text centered, draw=black}
}

\begin{tikzpicture}

\node (start) [epa, align=center] at (0cm,0cm) {EPA ECOTOX};
%\node (epa-ecotox) [ecotox, above of = epa-ecotox-clean, align=center, yshift = 5cm] {EPA ECOTOX\\raw data};
%\node (epa-ecotox-clean) [ecotox, align = center] at (-3cm,0cm) {EPA ECOTOX\\cleaned};

\node (epa-prepare) [script-r, align=center, right of=start, xshift=3cm] {EPA prepare};

\end{tikzpicture}
\end{document}