### 2020-01-27 article/tables/database-differences.tex

%% table that compares different ecotoxicological databases

%%%% CONTINUE HERE

\begin{tabular}{|m{3cm}|m{3cm}|m{1cm}|m{2cm}|m{1cm}|m{2cm}|l|}
\hline
Database & Publisher & Filter & Aggregation, Selection & Scriptable & Access & website \\
\hline
Comptox & Environmental Protection Agency & yes & no & no & web & https://comptox.epa.gov/dashboard \\
\hline
Ecotox & Environmental Protection Agency & yes & no & no & web & https://webetox.uba.de/webETOX/index.do \\
\hline
EnviroTox & Health and Environmental Sciences Institute & yes & per species & no & web & https://envirotoxdatabase.org \\
\hline
Etox & Umweltbundesamt & yes & no & no & web & https://webetox.uba.de/webETOX/index.do \\
\hline
Pesticide Property Data Base (PPDB) & University of Hertfordshire & fixed values & manual selection & no & web, file & https://sitem.herts.ac.uk/aeru/ppdb/index.htm \\
\hline
\textbf{Standartox} & \textbf{this article} & \textbf{yes} & \textbf{per chemical and organism} & \textbf{yes} & \textbf{web, API} & \textbf{http://standartox.uni-landau.de} \\
\hline
\end{tabular}