### 2020-01-27 article/tables/endpoint-conflate.tex

\begin{tabular}{|l|l|}
\hline
Standartox endpoint & Ecotox endpoint & Ecotox endpoint description \\ \hline
XX50 & LC50 & Lethal concentration to 50\% of test organisms \\ \hline
XX50 & LD50 & Lethal dose to 50\% of test organisms \\ \hline
XX50 & EC50 & Effective concentration to 50\% of test organisms \\ \hline
XX50 & ED50 & Effective dose to 50\% of test organisms \\ \hline
XX50 & IC50 & Inhibition concentration to 50\% of test organisms \\ \hline
XX50 & ID50 & Inhibition dose to 50\% of test organisms \\ \hline
XX50 & ET50 & Effective response time to 50\% of test organisms \\ \hline
XX50 & LT50 & Time to 50\% mortality of organisms \\ \hline
NOEX & NOEC & No-observable-effect-concentration \\ \hline
NOEX & NOEL & No-observable-effect-level \\ \hline
LOEX & LOEC & Lowest observable effect concentration\\ \hline
LOEX & LOEL & Lowest-observable-effect-level \\ \hline
\end{tabular}