### 2020-03-01 article/parts/10_supplement.tex

\section*{Supplement}

%% Endpoints conflate
\begin{table}
    \label{tab:endpoints-conflate}
    \catption{Table of Standartox endpoints and their corresponding Ecotox endpoints and a description.}
    \input{article/tables/endpoint-conflate.tex}
\end{table}

%% Additional databases
\begin{table}[ht]
    \label{tab:data-base-additional}
    \caption{Table of additionally queried publicly available data bases and their URLs.}
    \input{article/tables/data-base-tables.tex}
\end{table}

%% App parameters
(See Table \ref{tab:app-parameters})

\pagebreak
\section{CHECK THIS}

It can either be accessed through the APP (Figure \ref{fig:app}) or the API \ref{tab:api-endpoints} together with the R-package standartox (Code example \ref{listing:standartox-example}).

\subsection*{R-package list}
\label{list:r-packages}
\input{article/tables/r_package_list.tex}

\subsection*{Unit conversions}
 TODO
\begin{table}
    \csvautotabular[respect all]{lookup/lookup_concentration_unit.csv}
    \caption{Concentration unit conversion}
    \label{sup:conv-concentration}
\end{table}

\begin{table}
    \csvautotabular[respect all]{lookup/lookup_duration.csv}
    \caption{Duration unit conversion}
    \label{sup:conv-duration}
\end{table}

%\input{article/tables/chck-units-concentration.tex}

%\input{article/tables/chck-units-duration.tex}

\subsection*{Scripts}
\begin{figure}
    % TODO make better graphic
    \includegraphics[width=0.5\linewidth]{article/figures/scripts_dendrogram.png}
    \caption{Processing pipeline}
    \label{fig:pipeline-tree}
\end{figure}