### 2020-03-14 article/tikz/standartox-organigram.tex

\documentclass[border = 1.5cm]{standalone}
\usepackage{tikzpeople}
\usetikzlibrary{trees,
                arrows,
                shapes.geometric,
                positioning,
                calc,
                backgrounds,
                fit}
%% document
\begin{document}
%% layers
\pgfdeclarelayer{bg1}    % declare background layer
\pgfsetlayers{bg1, main}  % set the order of the layers (main is the standard layer)
%% styles
\tikzset{
	font = {\sffamily},
	ecotox/.style = {rectangle, minimum width = 4cm, minimum height = 3cm, text centered, rounded corners, draw = black, fill = yellow!10},
	standartox/.style = {rectangle, minimum width = 4cm, minimum height = 2cm, text centered, rounded corners, draw = black, fill = purple!10},
	app/.style = {rectangle, minimum width = 4cm, minimum height = 2cm, text centered, rounded corners, draw = black, fill = red!10},
	frame/.style = {draw, rectangle, rounded corners, dashed, blue, line width = 1pt, inner sep = 4mm},
	data/.style = {rectangle, minimum width = 7cm, minimum height = 3cm, text centered, rounded corners, draw = black},
	ids/.style = {diamond, minimum width = 3cm, minimum height = 2cm, text centered, draw = black, fill = green!10},
	db/.style = {rectangle, text centered, draw = black, rounded corners, minimum width = 3cm, minimum height = 1.25cm},
	arrow1/.style = {->, > = stealth, draw = black, color = gray!80, line width = 3pt}
}
\begin{tikzpicture}
%% processing
\node (epa-ecotox) [ecotox, above of = ecotox-id, align=center, xshift = -3cm, yshift = 4cm] {\Large EPA ECOTOX\\\Large raw data};
\node (ecotox-id) [standartox, align = center] at (-3cm,0cm) {\Large Identifiers};
\node (ecotox-unit) [standartox, below of = ecotox-id, align = center, yshift = -2cm] {\Large Unit\\\Large harmonisation};
\node (ecotox-chck) [standartox, below of = ecotox-unit, align = center, yshift = -2cm] {\Large Check\\\Large Conversions};
\node (cas) [ids, below of = ecotox-id, xshift = 6cm, yshift = 3cm] {\Large CAS};
\node (taxa) [ids, below of = cas, xshift = 0cm, yshift = -2.5cm] {\Large Taxa};
\node (db-chemical) [db, right of = cas, xshift = 4cm, align = center] {\Large Chemical\\\Large databases};
\node (db-habitat) [db, right of = taxa, xshift = 4cm, yshift = 1cm, align = center] {\Large Habitat\\\Large databases};
\node (db-region) [db, right of = taxa, xshift = 4cm, yshift = -1cm, align = center] {\Large Region\\\Large databases};
%% standartox
\node (stx-data) [standartox, below of = ecotox-chck, xshift = 3cm, yshift = -2.5cm, align = center] {\Large Standartox\\\Large data set};
\node (stx-filter) [standartox, below of = stx-data, xshift = 0cm, yshift = -2cm, align = center] {\Large Filter\\\Large methods};
\node (stx-agg) [standartox, below of = stx-filter, xshift = 0cm, yshift = -2cm, align = center] {\Large Aggregation\\\Large methods};
\node (stx-app) [app, below of = stx-data, yshift = 1cm, xshift = 6cm, align = center] {\Large Web\\\Large application};
\node (stx-rpackage) [app, below of = stx-data, yshift = -4cm, xshift = 6cm, minimum height = 5cm] {\Large R-package};
\node (stx-api) [app, below of = stx-rpackage, yshift = 2.25cm, xshift = 0cm, minimum width = 3cm, minimum height = 1.5cm] {\Large API};
%% arrows
\begin{pgfonlayer}{bg1}
    \draw [arrow1] (epa-ecotox.south) to [out=75,in=180] (stx-data.west);
    \draw [arrow1] (stx-data) -- (stx-filter);
    \draw [arrow1] (stx-filter) -- (stx-agg);
    \draw [arrow1] (ecotox-id.east) -- (cas.west);
    \draw [arrow1] (ecotox-id.east) -- (taxa.west);
    \draw [arrow1] (cas.east)-- (db-chemical.west);
    \draw [arrow1] (taxa.east) -- (db-habitat.west);
    \draw [arrow1] (taxa.east) -- (db-region.west);
\end{pgfonlayer}
%% frames
\node (frame-stx) [frame, fit = (stx-data) (stx-rpackage) (stx-app) ] {};
\node (frame-stx-text) [above, blue, inner sep = 2mm] at (frame-stx.north) {\Large Standartox};
%% people
\node (person) [bob, right of = stx-data, xshift = 9cm, yshift = -8cm, minimum size = 2cm] {};
\end{tikzpicture}
\end{document}