### 2020-05-04 Rmd/rep_conv_unit_result_duration.Rmd
release 0.1

---
title: "Unit conversion report"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      fig.width = 15, fig.height = 8)
```

```{r data, echo=FALSE, include=FALSE}
## converted units
q = "SELECT *
     FROM ecotox.results2"
units = read_query(user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
                   query = q)
## conversion check
chck_unit_result_conversion = readRDS(file.path(cachedir, 'chck_unit_result_conversion.rds'))
chck_unit_duration_conversion = readRDS(file.path(cachedir, 'chck_unit_duration_conversion.rds'))
```

## Result units

### Distinct units reduction
```{r}
res_n = units[ , lapply(.SD, uniqueN), .SDcols = c('conc1_unit', 'conc1_unit2', 'conc1_unit3', 'conc1_unit4')]
ggplot(melt(res_n), aes(y = value, x = variable)) +
  geom_bar(stat = 'identity')
```

Number of tests removed b/c of removed units: `r tab = table(units$conc1_remove); round(tab[2] / nrow(units) * 100, 1)`

### Number of tests per kept original units
Area is on `log10` scale.

```{r}
res_dist = units[ conc1_remove == FALSE, .N, conc1_unit ]
res_dist[ , label := paste0(conc1_unit, '\n', N) ]
ggplot(res_dist, aes(area = log10(N), label = label)) +
  geom_treemap() +
  geom_treemap_text(col = 'white')
```

### Number of tests per removed original units
Area is on `log10` scale.

```{r}
res_dist = units[ conc1_remove == TRUE, .N, conc1_unit ]
res_dist[ , label := paste0(conc1_unit, '\n', N) ]
ggplot(res_dist, aes(area = log10(N), label = label)) +
  geom_treemap() +
  geom_treemap_text(col = 'white')
```

### Number of tests per kept converted units
Area is on `log10` scale.

```{r}
res_dist = units[ conc1_remove == FALSE, .N, conc1_unit4 ]
res_dist[ , label := paste0(conc1_unit4, '\n', N) ]
ggplot(res_dist, aes(area = log10(N), label = label)) +
  geom_treemap() +
  geom_treemap_text(col = 'white')
```

### Number of tests per removed converted units
Area is on `log10` scale.

```{r}
res_dist_rm = units[ conc1_remove == TRUE, .N, conc1_unit4 ]
res_dist_rm[ , label := paste0(conc1_unit4, '\n', N) ]
ggplot(res_dist_rm, aes(area = log10(N), label = label)) +
  geom_treemap() +
  geom_treemap_text(col = 'white')
```

### Conversion check

#### All units
```{r}
chck_res = chck_unit_result_conversion[ , .N, .(chck_mean2, chck_unit2, chck_mean4, chck_unit4) ]
m = melt(chck_res, measure.vars = c('chck_mean2', 'chck_unit2', 'chck_mean4', 'chck_unit4'))
m[ , chck := fifelse(value == TRUE, 'correct', fifelse(value == FALSE, 'wrong', NA_character_)) ]
m[ is.na(chck), chck := 'no chck\nvalue' ]
# plot
ggplot(m, aes(y = N, x = variable, group = chck, fill = chck, label = N)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  geom_text(position = position_dodge(width = 0.9),
            hjust = 1.2, col = 'white', size = 4) +
  scale_fill_manual(values = c('springgreen1', 'salmon1', 'firebrick4')) +
    labs(title = 'Standartox unit conversion check',
         subtitle = 'All units',
         x = NULL) +
  coord_flip()
```

#### Only not removed units
```{r}
chck_res_nrm = chck_unit_result_conversion[ conc1_remove != TRUE,
                                            .N,
                                            .(chck_mean2, chck_unit2, chck_mean4, chck_unit4) ]
m = melt(chck_res_nrm,
         measure.vars = c('chck_mean2', 'chck_unit2', 'chck_mean4', 'chck_unit4'))
m[ , chck := fifelse(value == TRUE, 'correct', fifelse(value == FALSE, 'wrong', NA_character_)) ]
m[ is.na(chck), chck := 'no chck\nvalue' ]
# plot
ggplot(m, aes(y = N, x = variable, group = chck, fill = chck, label = N)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  geom_text(position = position_dodge(width = 0.9),
            hjust = 1.2, col = 'white', size = 4) +
  scale_fill_manual(values = c('springgreen1', 'salmon1', 'firebrick4')) +
  labs(title = 'Standartox unit conversion check',
       subtitle = 'Only kept units',
       x = NULL) +
  coord_flip()
```

```{r paper-review, eval=FALSE}
result_id_excl = chck_unit_result_conversion[ conc1_unit4 %in% c('g/l', 'g/g', 'g/m2', 'ppb', 'l/l', 'l/m2') ][ chck_unit2 == FALSE | is.na(chck_unit2), result_id ]
chck_paper = chck_unit_result_conversion[ conc1_unit4 %in% c('g/l', 'g/g', 'g/m2', 'ppb', 'l/l', 'l/m2') &
                                            !result_id %in% result_id_excl ,
                                          .N,
                                          .(chck_unit2, chck_unit4) ]
conc1_unit_ = chck_unit_result_conversion[ conc1_unit4 %in% c('g/l', 'g/g', 'g/m2', 'ppb', 'l/l', 'l/m2') &
                                             !result_id %in% result_id_excl,
                                           conc1_unit ]

q = "SELECT conc1_unit, count(*) n
 FROM ecotox.results2
 GROUP BY conc1_unit
 ORDER BY n DESC"
dat = read_query(user = DBuser, host = DBhost, port = DBport, password = DBpassword, dbname = DBetox,
                 query = q)
dat[ , kept := fifelse(conc1_unit %in% conc1_unit_, TRUE, FALSE) ]
dat2 = dat[ , .(sum(n)), kept]
dat2$V1[1] / sum(dat2$V1)

m = melt(chck_paper,
         measure.vars = c('chck_unit2', 'chck_unit4'))
m[ , chck := fifelse(value == TRUE, 'correct', fifelse(value == FALSE, 'wrong', NA_character_)) ]
m[ is.na(chck), chck := 'no chck\nvalue' ]
# plot
ggplot(m, aes(y = N, x = variable, group = chck, fill = chck, label = N)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  geom_text(position = position_dodge(width = 0.9),
            hjust = 1.2, col = 'white', size = 4) +
  scale_fill_manual(values = c('springgreen1', 'salmon1', 'firebrick4')) +
  labs(title = 'Standartox unit conversion check',
       subtitle = 'Only kept units',
       x = NULL) +
  coord_flip()
```

## Duration units

```{r}
dur_n = units[ , lapply(.SD, uniqueN), .SDcols = c('obs_duration_unit', 'obs_duration_unit2') ]
ggplot(melt(dur_n), aes(y = value, x = variable)) +
  geom_bar(stat = 'identity')
```

### Number of tests per kept original units
Area is on `log10` scale.

```{r}
dur_dist = units[ conc1_remove == FALSE, .N, obs_duration_unit ]
dur_dist[ , label := paste0(obs_duration_unit, '\n', N) ]
ggplot(dur_dist, aes(area = log10(N), label = label)) +
  geom_treemap() +
  geom_treemap_text(col = 'white')
```

### Number of tests per removed original units
Area is on `log10` scale.

```{r}
dur_dist = units[ conc1_remove == TRUE, .N, obs_duration_unit ]
dur_dist[ , label := paste0(obs_duration_unit, '\n', N) ]
ggplot(dur_dist, aes(area = log10(N), label = label)) +
  geom_treemap() +
  geom_treemap_text(col = 'white')
```

### Number of tests per kept converted units
Area is on `log10` scale.

```{r}
dur_dist = units[ conc1_remove == FALSE, .N, obs_duration_unit2 ]
dur_dist[ , label := paste0(obs_duration_unit2, '\n', N) ]
ggplot(dur_dist, aes(area = log10(N), label = label)) +
  geom_treemap() +
  geom_treemap_text(col = 'white')
```

### Number of tests per removed converted units
Area is on `log10` scale.

```{r}
dur_dist_rm = units[ conc1_remove == TRUE, .N, obs_duration_unit2 ]
dur_dist_rm[ , label := paste0(obs_duration_unit2, '\n', N) ]
ggplot(dur_dist_rm, aes(area = log10(N), label = label)) +
  geom_treemap() +
  geom_treemap_text(col = 'white')
```

### Conversion check

#### All units
```{r}
chck_dur = chck_unit_duration_conversion[ , .N, .(chck_mean2, chck_unit2) ]
m = melt(chck_dur, measure.vars = c('chck_mean2', 'chck_unit2'))
m[ , chck := fifelse(value == TRUE, 'correct', fifelse(value == FALSE, 'wrong', NA_character_)) ]
m[ is.na(chck), chck := 'no chck\nvalue' ]
# plot
ggplot(m, aes(y = N, x = variable, fill = chck, label = N)) +
  geom_bar(stat = 'identity', position = position_dodge()) +
  geom_text(position = position_dodge(width = 0.9), hjust = 1.2, col = 'white', size = 4) +
  scale_fill_manual(values = c('springgreen1', 'salmon1', 'firebrick4')) +
  labs(x = NULL) +
  coord_flip()
```

```{r plot, eval=FALSE, fig.width=10, fig.height=30}
# TODO plot is way to large
# ## 1
# test = units[ , .(n = .N), conc1_unit4 ]
# ggplot(test, aes(area = n, label = conc1_unit4)) +
#   geom_treemap() +
#   geom_treemap_text(col = 'white')
# 
# 
# # ## 2
# require(ggalluvial)
# 
# dat = units[ conc1_remove == TRUE, .N, .(conc1_unit, conc1_unit2, conc1_unit3, conc1_unit4, conc1_remove) ]
# setorder(dat, -N)
# 
# gg_alluvium = ggplot(dat,
#        aes(axis1 = reorder(conc1_unit, -N),
#            axis2 = reorder(conc1_unit2, -N), # axis3 = conc1_unit3, axis4 = conc1_unit4,
#            y = N, fill = conc1_remove)) +
#   geom_alluvium() +
#   geom_stratum(width = 1/12, fill = "black", color = "grey") +
#   geom_label(stat = "stratum", infer.label = TRUE)
#   
# ggsave(gg_alluvium, filename = file.path('/tmp/alluvium.svg'),
#        width = 20, height = 60, dpi = 90, limitsize = FALSE)

```